### 2013-03-13 text/opencv-test.tex
Added some youtube links to report

\documentclass{article}
\usepackage{graphicx}
\usepackage{hyperref}

\begin{document}

\title{Preliminary test on OpenCV feature implementations}
\author{Jonas Arnfred}

\maketitle

\begin{abstract}
% TODO!
\end{abstract}

\section{Project Description} 
In controlled settings existing face recognition
algorithms have achieved good results in both classifying and clustering faces,
but often these approaches break down when we consider images from real life 
application such as cell phone photo albums. Here we encounter non-aligned 
faces under different lighting conditions that might be partially obscured and 
turning in different directions. Even recent approaches to recognize 
non-aligned faces such as \cite[pp 9.]{liao2011} don't achieve more than 60\% 
geniune accept rate at 1\% false accept rate when comparing faces from 
LFW\footnote{Labeled Faces in the Wild} database. When we consider face patches 
not revealing the entire face the performance drops further to around 40\% 
genuine accept rate at 1\% false accept rate.

If we relax the problem of classification to the more general problem of 
clustering it might turn out that this accuracity is sufficient to accurately 
separate a set of face patches into partitions each ideally only containing 
pictures belonging to one person.

\subsection{A modest proposal}
%
% Consider a graph $G = (V, E)$ where $v_i$ is a vertix if $v_i \in V$ and 
% $e_i$ is an edge if $(v_i, v_j) \in E$ for $v_i, v_j \in V$ and a class 
% function $C (V \rightarrow c (c \in C)$ it is often possible to obtain a 
% partitioning $P$ such that for each partition $p \in P$ we have that $\forall 
% v_i, v_j \in c, C(v_i) = C(v_j)$ \emph{even} when the graph contains 
% interclass edges, that is an $\exists e \in E$ for which $
%
If we consider a graph it is often possible to succesfully obtain a 
partitioning that adheres to a classification of the graph even if the graph 
has intercluster edges. This is of course only possible as long as the 
intercluster edges outnumber or outweigh the intracluster edges, but if we can 
create a link between two nodes around 40\% of the time and only have 1\% of 
them be incorrect, we are looking at a very well behaved graph.

To apply graph clustering to image clustering, consider the set of image 
patches as the set of nodes and the edges between the nodes as the similarity 
between two image patches. The ideal partitioning of the graph is then one 
where all the nodes representing image patches belonging to a particular person 
are clustered together in their own partition.

Intuitively this is an easier problem than classification. Even if two 
images-patches might not resemble each other, they could both happen to 
resemble a common neighbor and consequently be clustered together.

\subsection{How can this be done?}
Because images in real world conditions are often taken under changing light 
and containing partially obscured or half turned faces, we can't require images 
to be aligned and homogenious in lighting and position in order to use them.

An alternative approach would be to sample characteristics from the face-patch 
in the form of feature points and compare how well these samples match feature 
points from other face-patches. If we can describe the feature points in a 
light and affine transformation agnostic way, ideally we would be able to match 
them to corresponding feature points on other face-patches.

By comparing feature points we can construct a similarity graph linking the 
face-patches to each other and use a graph clustering algorithm to obtain the 
final clusterings.

\section{So is it done yet?}
Before I present the finished algorithm there are a few practical and 
theoretical problems that need to be solved first.

\subsection{Theoretical shenanigans\footnote{%
\href{http://www.youtube.com/watch?v=MFZG8KQJni8}{Relevant YouTube Clip}}}

Given two sets of feature points and trying to match them, it's important to 
realize that these two sets might be obtained from face-patches that contain 
some overlapping characteristics but both might also contain features that are 
not present in the other patch at all. If one patch contain the face from the 
eyes and down (the person might be wearing a cap) and the other patch contain 
everything from the nose and up (the same person might cover their mouth with 
their hand) we will have feature points from the cap, the hair, the mouth and 
the hand that will not match at all. This means that when we try to match two 
face-patches, we need to match them in a way that highlights the points that 
match well and ignores the points that don't.

I see two ways to possibly approach the problem. One is to find a mathematical 
method like sub-tree matching or similar approaches to match the two sets of 
feature points against each other for each image\footnote{I would be interested 
in hearing your input on this subject}.  A different (and in my opinion 
promising) approach is to forego the matching of two images entirely:

If we consider the entire set of face-patches and the even bigger set of all 
the feature points, we could potentially match each feature point against the 
set of all possible feature points. Instead of basing the edges in the graph on 
the similarity of two image patches, we base the edge on the feature points 
that happen to be linked between any two given face-patches when matched 
against all possible points. A na\"ive approach is obviously of quadratic 
complexity in the amount of feature points, which very fast would become 
unfeasible to compute, but since points that can be compared both euclidean 
distances and hamming distances form a metric space, a much faster solution 
might be possible using metric trees like VP and BK trees\footnote{See: 
\href{http://stackoverflow.com/questions/6389841/efficiently-find-binary-strings-with-low-hamming-distance-in-large-set}{This 
answe on stackoverflow.com} and 
\href{http://en.wikipedia.org/wiki/Metric_tree}{Wikipedia} (I promise and swear 
I won't use wikipedia as a source for the thesis!)
}. In the case of hamming distance I could potentially further Huffman encode 
the library of feature points to reduce their length without reducing 
dimensionality.

\subsection{Practical considerations}

The success of this approach is dependent on feature points being able to 
describe image keypoints in a light and affine transformation agnostic way. In 
the paper using an alignment free approach to match images I mentioned 
before\footnote{\cite{liao2011}} they whip up their own feature descriptor, an 
approach I'm not personally very convinced of since they never show their own 
descriptor to be superior to anything else than sift. Given that many new 
feature descriptors have come out in recent years, it seems to me much more 
sensible to use an already tested and confirmed approach than to cook up a new 
method, before I can show that coming up with a new feature descriptor is the 
only way to solve the problem\footnote{And in this case I'll most likely spend 
my thesis trying to come up with a feature descriptor instead of solving the 
original problem}.

Fortunately OpenCV implements a range of feature descriptors\footnote{To be 
precise they include an interface for: SIFT, SURF, ORB, BRISK, BRIEF and FREAK.  
I have however not been able to get FREAK to work}. So in order to see if any 
of these feature detectors might be usable and if yes, which ones, I decided to 
put them to the test.

\section{Getting some data\footnote{%
\href{http://www.youtube.com/watch?v=FQ3UrLQP2ok}{Relevant YouTube clip}}}

Here is the text of your introduction.

\begin{equation}
    \label{simple_equation}
    \alpha = \sqrt{ \beta }
\end{equation}

\subsection{Subsection Heading Here}
Write your subsection text here.

\begin{figure}
    \centering
    \includegraphics[width=3.0in]{myfigure}
    \caption{Simulation Results}
    \label{simulationfigure}
\end{figure}

\section{Conclusion}
Write your conclusion here.

\end{document}