### 2013-03-13 text/opencv-test.tex
Finished the opencv-test writeup

\documentclass{article}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}

\begin{document}

\title{Preliminary test on OpenCV feature implementations}
\author{Jonas Arnfred}

\maketitle

\begin{abstract}
% TODO!
\end{abstract}

\section{Project Description} 
In controlled settings existing face recognition
algorithms have achieved good results in both classifying and clustering faces,
but often these approaches break down when we consider images from real life 
application such as cell phone photo albums. Here we encounter non-aligned 
faces under different lighting conditions that might be partially obscured and 
turning in different directions. Even recent approaches to recognize 
non-aligned faces such as \cite[pp 9.]{liao2011} don't achieve more than 60\% 
geniune accept rate at 1\% false accept rate when comparing faces from 
LFW\footnote{Labeled Faces in the Wild} database. When we consider face patches 
not revealing the entire face the performance drops further to around 40\% 
genuine accept rate at 1\% false accept rate.

If we relax the problem of classification to the more general problem of 
clustering it might turn out that this accuracy is sufficient to accurately 
separate a set of face patches into partitions each ideally only containing 
pictures belonging to one person.

\subsection{A modest proposal}
%
% Consider a graph $G = (V, E)$ where $v_i$ is a vertix if $v_i \in V$ and 
% $e_i$ is an edge if $(v_i, v_j) \in E$ for $v_i, v_j \in V$ and a class 
% function $C (V \rightarrow c (c \in C)$ it is often possible to obtain a 
% partitioning $P$ such that for each partition $p \in P$ we have that $\forall 
% v_i, v_j \in c, C(v_i) = C(v_j)$ \emph{even} when the graph contains 
% interclass edges, that is an $\exists e \in E$ for which $
%
If we consider a graph it is often possible to successfully obtain a 
partitioning that adheres to a classification of the graph even if the graph 
has \emph{inter}-cluster edges. This is of course only possible as long as the 
\emph{inter}-cluster edges outnumber or outweigh the \emph{intra}-cluster 
edges, but if we can create a link between two nodes around 40\% of the time 
and only have 1\% of them be incorrect, we are looking at a very well behaved 
graph.

To apply graph clustering to image clustering, consider the set of image 
patches as the set of nodes and the edges between the nodes as the similarity 
between two image patches. The ideal partitioning of the graph is then one 
where all the nodes representing image patches belonging to a particular person 
are clustered together in their own partition.

Intuitively this is an easier problem than classification. Even if two 
images-patches might not resemble each other, they could both happen to 
resemble a common neighbor and consequently be clustered together.

\subsection{How can this be done?}
Because images in real world conditions are often taken under changing light 
and containing partially obscured or half turned faces, we can't require images 
to be aligned and homogenious in lighting and position in order to use them.

An alternative approach would be to sample characteristics from the face-patch 
in the form of feature points and compare how well these samples match feature 
points from other face-patches. If we can describe the feature points in a 
light and affine transformation agnostic way, ideally we would be able to match 
them to corresponding feature points on other face-patches.

By comparing feature points we can construct a similarity graph linking the 
face-patches to each other and use a graph clustering algorithm to obtain the 
final clusterings.

\section{So is it done yet?}
Before I present the finished algorithm there are a few practical and 
theoretical problems that need to be solved first.

\subsection{Theoretical shenanigans\footnote{%
\href{http://www.youtube.com/watch?v=MFZG8KQJni8}{http://www.youtube.com/watch?v=MFZG8KQJni8}}}

Given two sets of feature points and trying to match them, it's important to 
realize that these two sets might be obtained from face-patches that contain 
some overlapping characteristics but both might also contain features that are 
not present in the other patch at all. If one patch contain the face from the 
eyes and down (the person might be wearing a cap) and the other patch contain 
everything from the nose and up (the same person might cover their mouth with 
their hand) we will have feature points from the cap, the hair, the mouth and 
the hand that will not match at all. This means that when we try to match two 
face-patches, we need to match them in a way that highlights the points that 
match well and ignores the points that don't.

I see two ways to possibly approach the problem. One is to find a mathematical 
method like sub-tree matching or similar approaches to match the two sets of 
feature points against each other for each image\footnote{I would be interested 
in hearing your input on this subject}.  A different (and in my opinion 
promising) approach is to forego the matching of two images entirely:

If we consider the entire set of face-patches and the even bigger set of all 
the feature points, we could potentially match each feature point against the 
set of all possible feature points. Instead of basing the edges in the graph on 
the similarity of two image patches, we base the edge on the feature points 
that happen to be linked between any two given face-patches when matched 
against all possible points. A na\"ive approach is obviously of quadratic 
complexity in the amount of feature points, which very fast would become 
unfeasible to compute, but since points that can be compared both euclidean 
distances and hamming distances form a metric space, a much faster solution 
might be possible using metric trees like VP and BK trees\footnote{See: 
\href{http://stackoverflow.com/questions/6389841/efficiently-find-binary-strings-with-low-hamming-distance-in-large-set}{This 
answe on stackoverflow.com} and 
\href{http://en.wikipedia.org/wiki/Metric_tree}{Wikipedia} (I promise and swear 
I won't use wikipedia as a source for the thesis!)
}. In the case of hamming distance I could potentially further Huffman encode 
the library of feature points to reduce their length without reducing 
dimensionality.

\subsection{Practical considerations}

The success of this approach is dependent on feature points being able to 
describe image keypoints in a light and affine transformation agnostic way. In 
the paper using an alignment free approach to match images I mentioned 
before\footnote{\cite{liao2011}} they whip up their own feature descriptor, an 
approach I'm not personally very convinced of since they never show their own 
descriptor to be superior to anything else than sift. Given that many new 
feature descriptors have come out in recent years, it seems to me much more 
sensible to use an already tested and confirmed approach than to cook up a new 
method, before I can show that coming up with a new feature descriptor is the 
only way to solve the problem\footnote{And in this case I'll most likely spend 
my thesis trying to come up with a feature descriptor instead of solving the 
original problem}.

Fortunately OpenCV implements a range of feature descriptors\footnote{To be 
precise they include an interface for: SIFT, SURF, ORB, BRISK, BRIEF and FREAK.  
I have however not been able to get FREAK to work}. So in order to see if any 
of these feature detectors might be usable and if yes, which ones, I decided to 
put them to the test.

\section{Experiments}
The goal of these experiment is to see if different feature descriptors given 
real life images would score the similarities between images showing the same 
face significantly higher than images showing the faces of different people.

\subsection{Getting some data\footnote{%
\href{http://www.youtube.com/watch?v=FQ3UrLQP2ok}{http://www.youtube.com/watch?v=FQ3UrLQP2ok}}}
In order to test this I used the LFW dataset\footnote{Labeled Faces in the 
Wild}.  This is a dataset consisting of labeled photos of famous people taken 
by photographers under different circumstances and in different quality. The 
faces vary between looking straight at the camera and turning their head to 
either side, but not to the extend of being full profile. The lighting is 
usually relatively good but very inconsistent as most photos are from different 
events, sometimes years apart.

This dataset is not a perfect match since the ideal application for this 
algorithm would be people's own photo albums, where the camera used is 
consistent and the faces often might be in profile. However it does pose 
similar challenges in the form of lightning and non-aligned faces that it 
serves well to give a rough idea of the general performance of different 
feature descriptors.

\subsection{Experiment Setup}

In general there are two measures to consider when we compare keypoints in an 
image; the distance measure and how unique the match is. The distance measure 
for the keypoint descriptors I have reviewed are either based on euclidean 
distance or on hamming distance. Once we have matched keypoints of one image to 
their highest matching counterpart in another image, we can filter these 
matches either by only including matches that symmetric, i.e. if we swap the 
two images and run the same matching algorithm we get the same match. Another 
method proposed in \cite{lowe2004sift} is to look at the ratio between the best 
and second best match. If the distances are too similar, the keypoint match is 
not unique enough and hence discarded.

Since I don't have a solution yet for how best to compare two face-patches, I 
have created four na\"ive methods that gives a score for a match of two 
face-patches given their keypoint descriptors. These methods might not be 
ideal, but it's my hope that they serve the purpose of rating the similarities 
between images well enough to evaluate the different types of descriptors.  
Notice that since a better matching function would be able to discard more bad 
matches and keep more good matches, the results obtained using my na\"ive 
methods sets a lower bound on how well the different descriptors perform.  
Hopefully this lower bound is equally tight for all of them, but this is by no 
means a given and my comparison can therefore not be seen as a general 
evaluation of their performance.

The four matching functions are given as parameters the distance for each 
keypoint to it's closest match and the ratio for each keypoint between the 
distance of the closest and second closest match. They work as follows:
\begin{itemize}
	\item{\emph{Simple Mean}: Take the mean of all distances}
	\item{\emph{Better half}: Take the mean of the best half of the distances}
	\item{\emph{Treshold}: Take the mean of only the points that are both in the 
	better half of distances and ratios}
	\item{\emph{Limit to 10}: As above but take only the 10 points with the 
best distance}
\end{itemize}
I also created another four similarly functions, the only difference being that 
the mean was over the \emph{ratio} and not the \emph{distance}.

To quickly weed out bad performers I designed a small test-set including three 
people from the LFW testset\footnote{Anna Kournikova, Ann Veneman and Bill 
Gates}. For each person 10 images were picked at random and all pairs of images 
compared using the eight functions mentioned above.

\begin{table}[t]
\begin{center}
	\begin{tabular}{l l|l l}
		\multicolumn{2}{c}{Preliminary Test} & \multicolumn{2}{c}{Full Test} \\
		\hline
		Locator & Descriptor & Locator & Descriptor \\
		\hline
		Sift & Sift & &\\
		Surf & Surf & &\\
		Orb & Orb & Orb & Orb \\
		Fast & Orb & Fast & Orb \\
		Orb & Brief & Orb & Brief \\
		Fast & Brief & Fast & Brief \\
		Orb & Brisk & & \\
		Brisk & Brisk & & \\
	\end{tabular}
	\label{table:features}
	\caption{Keypoints and descriptors used in preliminary and full testing 
round}
\end{center}
\end{table}

Since I have several different methods for selecting keypoints and several 
different algorithms for creating descriptors based on the keypoints 
available in OpenCV. The pairings I tested can be seen in table 1.  
Generally I used the keypoint locator that was available for the 
descriptor I was testing, and when this wasn't possible paired the 
descriptor with a keypoint locator of a similar method.

Upon finding the best performing keypoint locator and descriptor pairs I went 
on to test them on an imageset containing $60$ people with ten images per 
person, that is $600$ images in total.

\section{Results\footnote{%
\href{http://www.youtube.com/watch?v=tKT-eWMWXOE}{http://www.youtube.com/watch?v=tKT-eWMWXOE}}}

The results from the preliminary testing can be seen in table 2.  The 
difference displayed is the percentage wise difference between the score 
of pairs of faces belonging to the same person as opposed to pairs of 
faces belonging to two different people. The score is calculated using 
either the ratio or the distance with one of the four functions 
described above.

It is immediately obvious that the distance based measures fare much better 
than the ratio based measures. In particular the `treshold' and `limit10' 
methods work well. However our main concern is not with how well the different 
functions perform but rather the performance of different descriptors. Here we 
see a clear advantage of Orb, Brief and Brisk. Brisk however shows some strange 
fluctuations and would be surprised if these results were a true indicator of 
the performance of the descriptor. Brisk is also significantly slower than the 
others, usually taking around 9-10 seconds to calculate similarities between 30 
images, something than can be calculated by the other methods in around 0.3 
seconds.

\begin{table}[t]
\begin{adjustbox}{center}
	\label{table:preliminary}
	\begin{tabular}{l c c c c c c c c}
		& \multicolumn{4}{c}{Ratio} & \multicolumn{4}{c}{Distance} \\
		& Mean & Best Half & Treshold & Limit10 & Mean & Best Half & Treshold &
Limit10 \\
		\hline
		Sift-Sift	& 0.4 & 0.7 & 0.8  & 0.4  & 0.7 & -0.4 & 1.0 & 1.5 \\
		Surf-Surf	& 0.4 & 0.7 & 0.8  & 1.4  & 0.9 & 0.9  & 0.7 & 0.9 \\
		Orb-Orb		& 0.6 & 1.1 & 1.3  & 1.1  & 2.0 & 0.6  & 3.8 & 3.5 \\
		Fast-Orb	& 0.8 & 1.5 & 1.7  & 1.6  & 2.7 & 13.2 & 4.3 & 4.7 \\
		Orb-Brief	& 0.4 & 0.9 & 1.0  & 0.6  & 5.8 & 2.3  & 8.9 & 8.8 \\
		Fast-Brief	& 1.1 & 2.0 & 2.2  & 2.0  & 4.8 & 11.0 & 8.3 & 10.6 \\
		Orb-Brisk	& 1.1 & 2.0 & 2.2  & 2.3  & 4.8 & 7.3  & 6.7 & 6.8 \\
		Brisk-Brisk	& 0.0 & 0.0 & -0.1 & -0.1 & 0.9 & -6.2 & 1.7 & 1.7 \\
	\end{tabular}
\end{adjustbox}
	\caption{Percentage difference in preliminary testing round}
\end{table}

The difference in mean doesn't tell us much if we don't know the variance of 
the individual classifier, since a high variance would mean that we would need 
an equally big difference to use the similarity score for anything. To 
illustrate the behavior of the distributions I've made a few plots. In 
figure \ref{fig:smallset-all-methods} eight histograms are shown, each 
displaying the behavior of a particular keypoint locator-descriptor 
combination using the same score function (Treshold).  When taking the 
distributions into account it's apparent that the descriptors Orb and 
Brief are performing about equally well given Orb's lower variance.

\begin{figure}
        \centering
		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/sift-sift-treshold}
				\caption{Sift-Sift Treshold: 1.0\% diff.}
				\label{fig:smallset-sift-sift-treshold}
        \end{subfigure}%
		\quad %add desired spacing between images, e. g. ~, \quad, \qquad etc.
          %(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/surf-surf-treshold}
				\caption{Surf-Surf Treshold: 0.6\% diff.}
				\label{fig:smallset-surf-surf-treshold}
        \end{subfigure}%

		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/orb-orb-treshold}
				\caption{Orb-Orb Treshold: 3.8\% diff.}
				\label{fig:smallset-orb-orb-treshold}
        \end{subfigure}%
		\quad%
		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/fast-orb-treshold}
				\caption{Fast-Orb Treshold: 4.3\% diff.}
				\label{fig:smallset-fast-orb-treshold}
        \end{subfigure}%

		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/orb-brief-treshold}
				\caption{Orb-Brief Treshold: 8.9\% diff.}
				\label{fig:smallset-orb-brief-treshold}
        \end{subfigure}%
		\quad%
		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/fast-brief-treshold}
				\caption{Fast-Brief Treshold: 8.3\% diff.}
				\label{fig:smallset-fast-brief-treshold}
        \end{subfigure}%

		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/orb-brisk-treshold}
				\caption{Orb-Brisk Treshold: 6.7\% diff.}
				\label{fig:smallset-orb-brisk-treshold}
        \end{subfigure}%
		\quad%
		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/brisk-brisk-treshold}
				\caption{Brisk-Brisk Treshold: 1.7\% diff.}
				\label{fig:smallset-brisk-brisk-treshold}
        \end{subfigure}%
		\caption{Treshold results on 30 images}\label{fig:smallset-all-methods}
\end{figure}

I've included figure \ref{fig:smallset-orb-brief} to illustrate how Orb-Brief 
performs over different score functions using the small testset of 30 images.  
With only 30 images, there is no saying if the scores are just a fluke, or if 
they will behave consistently at bigger datasets. Figure \ref{fig:bigset} is 
included to illustrate the behavior of Orb-Brief when testing on 600 images and 
60 people instead. The results are encouraging since they replicate and 
usual exceeds the preliminary results using only 30 images. In table 3 
the results from the simulation on 600 images are displayed.

\begin{figure}
        \centering
		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/orb-brief-simplemean}
				\caption{Orb-Brief Simple mean: 5.8\% diff.}
				\label{fig:smallset-orb-brief-simplemean}
        \end{subfigure}%
		\quad %add desired spacing between images, e. g. ~, \quad, \qquad etc.
          %(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[b]{0.45\textwidth}
				\centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/orb-brief-besthalf}
				\caption{Orb-Brief Best half: 2.3\% diff.}
				\label{fig:smallset-orb-brief-besthalf}
		\end{subfigure}

		\begin{subfigure}[b]{0.45\textwidth}
				\centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/orb-brief-treshold}
				\caption{Orb-Brief Treshold: 8.7\% diff.}
				\label{fig:smallset-orb-brief-treshold}
		\end{subfigure}
		\quad%
		\begin{subfigure}[b]{0.45\textwidth}
				\centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Smallset/orb-brief-limit10}
				\caption{Orb-Brief Limit 10: 8.8\% diff.}
				\label{fig:smallset-orb-brief-limit10}
		\end{subfigure}
		\caption{Orb-Brief results on 30 images}\label{fig:smallset-orb-brief}
\end{figure}

\begin{figure}
        \centering
		\begin{subfigure}[b]{0.45\textwidth}
                \centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Bigset/orb-brief-simplemean}
				\caption{Orb-Brief Simple mean: 6.3\% diff.}
				\label{fig:bigset-orb-brief-simplemean}
        \end{subfigure}%
        ~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
          %(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[b]{0.45\textwidth}
				\centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Bigset/orb-brief-besthalf}
				\caption{Orb-Brief Best half: 9.2\% diff.}
				\label{fig:bigset-orb-brief-besthalf}
		\end{subfigure}

		\begin{subfigure}[b]{0.45\textwidth}
				\centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Bigset/orb-brief-treshold}
				\caption{Orb-Brief Treshold: 9.3\% diff.}
				\label{fig:bigset-orb-brief-treshold}
		\end{subfigure}
%
		\begin{subfigure}[b]{0.45\textwidth}
				\centering
				\includegraphics[width=\textwidth]{../results/OpenCV-Evaluate/Bigset/orb-brief-limit10}
				\caption{Orb-Brief Limit 10: 9.8\% diff.}
				\label{fig:bigset-orb-brief-limit10}
		\end{subfigure}
		\caption{Orb-Brief results on 600 images}\label{fig:bigset}
\end{figure}

\begin{table}[t]
	\centering
	\label{table:bigset}
	\begin{tabular}{l c c c c}
		& Mean & Best Half & Treshold & Limit10 \\
		\hline
		Orb-Orb		& 2.4 & 3.8 & 4.1  & 4.4 \\
		Fast-Orb	& 3.9 & 5.7 & 6.2  & 7.1 \\
		Orb-Brief	& 6.3 & 9.2 & 9.3  & 9.8 \\
		Fast-Brief	& 4.7 & 7.0 & 7.3  & 8.2 \\
	\end{tabular}
	\caption{Percentage difference in full testing round}
\end{table}

\section{Conclusion and Questions for tomorrow}

In conclusion I think that both Orb and Brief would be worth testing out for an 
alignment free clustering approach. The important details I think we should 
touch upon tomorrow are as follows:
\begin{itemize}
\item{Do these results overall show enough discriminative power to merit 
further exploration?}
\item{How can I best compare the keypoints of two face-patches?}
\item{Does my proposal for using Metric trees to circumvent the problem sound 
doable?}
\item{Is there something super important that I've completely ignored so far?  
(I'm mostly sitting alone with this, so I need a reality check)}
\end{itemize}

\bibliographystyle{plainnat}
\bibliography{bibliography}

% Here is the text of your introduction.
% 
% \begin{equation}
%     \label{simple_equation}
%     \alpha = \sqrt{ \beta }
% \end{equation}
% 
% \subsection{Subsection Heading Here}
% Write your subsection text here.
% 
% \begin{figure}
%     \centering
%     \includegraphics[width=3.0in]{myfigure}
%     \caption{Simulation Results}
%     \label{simulationfigure}
% \end{figure}
% 
% \section{Conclusion}
% Write your conclusion here.

\end{document}