### 2013-08-09 text/head/titlepage.tex

\begin{titlepage}
\begin{center}
%\large
\sffamily

\null\vspace{2cm}
{\huge Reliable feature point matching} \\[24pt] 
\textcolor{gray}{\small{THIS IS A TEMPORARY TITLE PAGE \\ It will be 
replaced for the final print by a version \\ provided by the service 
academique.}}

\vfill

\begin{tabular} {cc}
\parbox{0.3\textwidth}{\includegraphics[width=4cm]{images/epfl}}
&
\parbox{0.7\textwidth}{%
	Thesis in Computer Science\\
	presented on September 3rd 2013\\
	at the Advanced Science Research Center\\
	Fusionopolis, Singapore\\

%
Singapore, ADSC, 2011}
\end{tabular}
\end{center}
\vspace{2cm}
\end{titlepage}