### 2013-08-09 text/thesis.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%		My Thesis
%
%		EDOC Template
%		2011
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{head/settings_epfl_template.tex}
\input{head/settings_custom.tex}  % place your custom packages, etc... in this file!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% HEAD: Book-Begin
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\frontmatter
\input{head/titlepage.tex}
%\include{head/dedication}
%\input{head/acknowledgements}
%\include{head/preface}
%\include{head/abstracts}

\tableofcontents
\addcontentsline{toc}{chapter}{List of figures} % adds an entry to the table of contents
\listoffigures
\addcontentsline{toc}{chapter}{List of tables} % adds an entry to the table of contents
\listoftables
% your list of symbols here, if needed.

% space before each new paragraph according to the template guidelines.
%(needs to be after titlepage and frontmatter to keep the table of contents lists short)
\setlength{\parskip}{1em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% MAIN: The chapters of the thesis
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\mainmatter
\include{main/ch1_introduction}
\include{main/ch2_background}
\include{main/ch3_matching_methods}
\include{main/ch4_competition}
\include{main/ch4_experiments}
\include{main/ch4_applications}
\include{main/ch4_summary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% TAIL: Bibliography, Appendix, CV
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\include{tail/appendix}
\include{tail/biblio}
% Add your glossary here
% Add your index here
% Photographic credits (list of pictures&images that have been used with names of the person holding the copyright for them)
% \include{tail/cv}

\end{document}