### 2013-08-10 text/main/ch2_background.tex
Added computer vision background

\chapter{Matching local image features}

In this chapter we will introduce local feature points in images and 
explain how they are used. We will start out by discussing why we need 
local features in the first place before looking at the inner workings 
of modern local image features. Finally we will look at how they are 
used, in particular how corresponding pairs are found in between two or 
more images.

\section{A brief look at computer vision}

The field of computer vision is anecdotally said to have been founded 
when Marvin Minsky assigned the problem of making a computer see as an 
undergraduate summer project back in 1966\footnote{This is the same 
Marvin who invented neural networks and went on to serve as scientific 
adviser on Kubrick's movie \emph{Space Odyssey 2001}}. Whether or not 
the anecdote is true, the idea that the field of computer vision should 
be solvable in a summer's worth of time but still to this days remains 
an active field of research, illustrates an interesting notion: Making 
computers solve tasks involving seeing turns out to be much harder than 
we would initially assume.

Part of this notion stems from the fact that vision is a problem that we 
as humans are very good at. We can easily recognize objects under 
different light, or stitch together photographies to a panorama, or 
navigate a corridor identifying obstacles along the way. In fact these 
things are so easy to us, that it seems strange that they should be 
difficult at all, yet consistently recognizing an object under different 
lighting conditions and from different angles can still cause problems 
for even the best object recognition algorithms.

The main difficulty lies in how the world we see around us constantly 
varies. Try for a moment to imagine that you enter a room and recognize 
the newspaper that is lying on the table, illuminated from the grey 
light coming in of the window on the opposing wall. You move across the 
room and switch on the light make out if the paper is from today, and 
then you sit down and put on your glasses to read the headline. Try to 
imagine the newspaper as you see it on three different stages. First as 
you enter the room, then as you turn after turning on the light, now 
standing on the other side of the table, and last when you are sitting 
with the paper in your hand after you put on your glasses. For each of 
these stages, picture the situation as a photography and look at how the 
newspaper changes as you move around. At first you see it at an angle 
lying in a fairly dark and grey room, the white pages reflecting the 
light from the window with a cold bluish hue and the words too far away 
to make out. Then in the second stage with light on, the object you are 
looking at has suddenly changed. When you turned on the light in the 
room, you also changed the color of the newspaper, which now has a warm 
and yellowy white color. The angle you are looking at the paper from has 
changed too, and the words and pictures are no longer upside down, 
although still slightly rotated and with the increased light in the room 
the contrast of the words and pictures of the page makes it possible for 
you to make recognize that the newspaper is from today. Finally as you 
sit down and put your glasses on the newspaper changes again. The color 
might be the same, but if you imagine taking a photography as you are 
sitting with the paper in your hand, the newspaper will look much bigger 
almost taking up the entire frame. Now that your glasses are on, the 
words stand out much clearer and you can see what is in the pictures on 
the front page.

Intuitively you know that the paper you sat down to read is the same 
object as the paper you saw when you entered the room, but if we limit 
ourselves to only looking at the object itself, almost everything about 
it changed in the process, from the color of the pages to the sharpness 
of the lines. On top of that the paper has changed in size and rotation 
from lying on the table to being in your hand. In our daily lives we are 
aided in the process of recognizing the paper by our memory and idea of 
how the world works. When we walked in to the dark room and saw 
something on the table, it was probably safe to assume that it was a 
newspaper. Maybe the newspaper always lies like that on the table, or 
maybe we left it there ourselves. Afterwards as we move around in the 
room, we can safely assume that the paper didn't move and even if it 
might not look exactly the same after we turned on the light, it's safe 
to guess that nobody took the old newspaper and replaced it with a new 
and yellower version.

On the other hand a computer algorithm trying to recognize objects will 
not be able to make the same assumptions about the world. We might just 
give as input to the program to pictures of the newspaper. One from the 
first stage where the light was turned off, the paper was far away and 
the lens wasn't focused correctly, and one from the third stage with the 
light on, the image in focus and the paper taking up the entire photo.  
Now we ask the algorithm: "Is this the same newspaper?". In this case it 
is fair to ask how the algorithm can reliably make this judgement.

\section{Introducing local features}

Introduction to Computer Vision
 - Introduce common tasks that we are good at that computers aren't
 - Single in on tasks that we can solve using feature points

Introduction to feature points
 - Global explanation of what they are and how they serve to solve the 
problems mentioned before
 - How do we find them? (Examples making it intuitive why corners are 
nice)
 - How do we describe them? (Relate invariance: Light, Rotation, Noise, 
 Scale, \ldots, Affine)
 - Introduce SIFT and explain how the different problems are tackled in 
SIFT)
 - Talk about other methods and how they are different

Matching feature points
 - What applications require matching?
 - How is matching usually done
 - Segue into related work from paper