### 2013-08-11 text/main/ch3_matching_methods.tex
Added some introduction to matching methods

\chapter{Algorithms}
\label{C:algorithms}

In this chapter we will describe the algorithms used in the experiements
in detail. We start with the proposed algorithms \emph{MM} and 
\emph{MMC}, and continue with the algorithms used for comparison, 
\emph{Ratio}, \emph{Spectral} and \emph{Isomatch}.

\section{Proposed Matching Methods}
\label{S:MatchingMethods}

\subsection{Mirror Match (\emph{MM})}

\begin{figure}[h]
	\centering%
		\begin{subfigure}[t]{0.5\columnwidth}
			\centering
			\includegraphics[width=0.95\columnwidth]{images/MMC_pitts_source}
			\caption{Source image pair}
			\label{fig:pitts_source}
		\end{subfigure}%
		~ %
		\begin{subfigure}[t]{0.5\columnwidth}
			\centering
			\includegraphics[width=0.95\columnwidth]{images/MMC_pitts_keypoints}
			\caption{Feature points}
			\label{fig:pitts_keypoints}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		%\qquad (or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.5\columnwidth}
			\centering
			\includegraphics[width=0.95\columnwidth]{images/mirror_match_off}
			\caption{\emph{Ratio}}
			\label{fig:unique}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, 
		\begin{subfigure}[t]{0.5\columnwidth}
			\centering
			\includegraphics[width=0.95\columnwidth]{images/mirror_match_with_pruned}
			\caption{\emph{MM} intermediate result}
			\label{fig:within}
		\end{subfigure}%
		\\ %add desired spacing between images, e. g. ~, \quad, 
		%\qquad (or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.5\columnwidth}
			\centering
			\includegraphics[width=0.95\columnwidth]{images/mirror_match}
			\caption{\emph{MM} final result}
			\label{fig:without}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, 
		\begin{subfigure}[t]{0.5\columnwidth}
			\centering
			\includegraphics[width=0.95\columnwidth]{images/MMC_partition}
			\caption{\emph{MMC} Partition Example}
			\label{fig:pitts_partition}
		\end{subfigure}%
	\caption{Feature matching with \emph{MM} and \emph{MMC}. Dots represent feature points; green/red lines indicate correct/incorrect matches, respectively; black lines represent edges in the feature graph.  
	(c) Result of \emph{Ratio} matching. 
  (d) All matches found by \emph{MM}, including intra-image matches. 
	(e) Final \emph{MM} result. 
	(f) Example of a partition of feature points after clustering, which 
includes similar feature points from building windows and shirt patterns.}%
	\label{fig:comparemirror}%
\end{figure}%

The central idea behind \emph{MM} is to match features of $n$ images by 
taking every feature from all $n$ images and matching it against every 
other feature from the same set. We can then discard the correspondences 
that match two points within the same image. Algorithm~\ref{alg-mm} 
details the implementation of \emph{MM}.

\begin{algorithm}[htb]
\caption{Mirror Match (\emph{MM})}
\label{alg-mm}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $images$ : set of images, $t \in \mathbb{R}$
\State $matches_{init}\gets \varnothing$
\State $matches_{final}\gets \varnothing$
\State $features\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Acquisition Stage
	\State $features\gets features \cup getFeatures(I_i)$
\EndFor
\ForAll{$f_i \in features$} \Comment Matching Stage
	\State $f_m,f_n \gets get2NNs(f_i, features \setminus 
	\left\{f_i\right\})$
	\State $ratio \gets similarity(f_i, f_n) / similarity(f_i, f_m)$
	\If{$ratio < t$}
		\State $matches_{init} \gets matches_{init} \cup \left(f_i, f_m\right)$
	\EndIf
\EndFor
\ForAll{$\left(f_i, f_j \right) \in matches_{init}$} \Comment Filter 
Stage
\If{$\left(f_j, f_i \right) \in matches_{init} \wedge getImg(f_i) \neq 
getImg(f_j) \wedge \left(f_j, f_i\right) \not\in matches_{final}$}
		\State $matches_{final} \gets (f_i, f_j)$
	\EndIf
\EndFor \\
\Return $matches_{final}$
\end{algorithmic}
%}
\end{algorithm}

In the acquisition stage we gather all features in the set of images.  
In the matching stage these features are matched using $k$-nearest 
neighbors.  For any given feature $f_i$ the two most similar neighbors 
are returned, and we calculate the ratio between them as proposed in 
\cite{lowe2004sift}.  Any correspondence with a ratio above the 
threshold supplied will be discarded. Finally in the filter stage we 
check that matches are from different images and discard all matches 
that are not symmetric.

Figure~\ref{fig:comparemirror} illustrates the benefits of \emph{MM} 
using an example image pair from the Gallagher dataset 
\cite{gallagher2008}.
With \emph{Ratio} (Figure~\ref{fig:unique}), many incorrect matches occur 
in the fence towards the bottom of the image.
When we match all feature points together, many of these incorrect 
matches are eliminated, because points in the fence match with other 
points in the fence in the same image (Figures~\ref{fig:within} and
\ref{fig:without}).

\subsection{Mirror Match with Clustering (\emph{MMC})}

In contrast to \emph{MM}, \emph{MMC} diverges from traditional 
non-geometric feature matching by clustering feature points by 
similarity. This process yields partitions of fairly similar feature 
points that we can match using the same approach as \emph{MM}.  
Algorithm~\ref{alg-mmc} shows the pseudocode implementation of \emph{MMC}.

Before we introduce the implementation details we will go over the 
problem of graph clustering and how it relates to feature matching.

To cluster the feature points, we construct a graph with the feature 
points as nodes and their similarity score as edge weights. Given a 
matching function $M(k_i, k_j) \rightarrow \mathbb{R}$ that takes two 
feature points in an image and returns their matching score, we can 
define a matrix $A$ where each element $A_ij = M(k_i, k_j)$. $A$ can be 
interpreted as the \emph{adjecency matrix} of the fully connected graph.

In the literature there are various ways of clustering a graph according 
to different measures of what constitutes an optimal partitioning. 
Traditionally the most used clustering algorithms have been K-means and 
spectral clustering, but in recent years a host of new algorithms have 
been proposed based on both Newman's concept of graph 
modularity\footnote{Introduced in \cite{girvan2002}, discussed in 
\cite{brandes2007} and used in \cite{blondel2008} as well as others} as 
well as information theoretical measures\footnote{See for example 
\cite{rosvall2008}} and the Potts spin model from physics\footnote{Used 
in \cite{ronhovde2009}} just to mention a few. On tests done using 
randomly generated graphs with a known partitioning \cite{blondel2008}, 
\cite{rosvall2008} and \cite{ronhovde2009} perform markedly better than 
spectral clustering and K-means\cite{lancichinetti2009}.

\begin{algorithm}[htb]
\caption{Mirror Match with Clustering (\emph{MMC})}
\label{alg-mmc}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $images$ : set of images, $t \in \mathbb{R}$
\State $M\gets \varnothing$
\State $F\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features
	\State $f_i\gets getFeatures(I_i)$
	\State $F\gets F \cup f_i$
\EndFor
\State $A\gets getAdjacencyMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $A_{norm}\gets 1 - normalize(A)$
\State $A_{pruned}\gets pruneEdges(A_{norm},\alpha)$
\State $P\gets cluster(A_{pruned})$
\ForAll{$p \in P$} \Comment p is a set of feature points
	\State $M\gets M \cup getMatches(p, t, F)$
\EndFor \\
\Return matches
\end{algorithmic}
%}
\end{algorithm}

\begin{figure}[t]
    \centering
	\includegraphics[width=\textwidth]{images/MMC_graph}
    \label{fig:graph}
	\caption{The partitioned feature graph. Each vertex represents a 
		feature point; lines indicate high similarity between points. A 
		partition is a connected group with the same color. The border 
		color of each node indicates which image it belongs to.  Zooming 
	into a subsection of the graph (right part), the various cluster 
sizes can be seen, ranging from hundreds of feature points to only two 
or three.}
	\label{fig:graph}
\end{figure}

We use the Louvain Method \cite{blondel2008} for clustering feature 
points, since it is relatively fast and performs well 
\cite{lancichinetti2009}, does not require parameters 
\cite{blondel2008}, and does not emphasize partitions of equal size, as 
opposed to spectral clustering or k-means \cite{von2007}, for example.
While the Louvain clustering algorithm does not require any parameters 
in itself, it tends towards clustering all feature points together in 
the same partition if the graph is very connected.  To ensure that the 
graph is well clustered, the adjacency matrix is pruned by setting all 
edges below a certain threshold to $0$.  From empirical analysis, 
retaining the top 2.5\% of edges with the highest similarity seems to 
work well in practice.  Figure~\ref{fig:graph} shows the result of 
clustering the feature points as a graph.

The partitions group feature points by similarity, which means that 
repetitive structures such as buildings often appear in larger 
partitions, as exemplified in Figure~\ref{fig:pitts_partition}.

\begin{algorithm}[htb]
\caption{Impl.\ of getMatches (\emph{from MMC algorithm})}
\label{alg-getmatches}
    \fontsize{10}{10}\selectfont
    \begin{algorithmic}
    \Require $p$ : set of features, $t\in \mathbb{R}$, $features$ : Set of 
    all features
    \State $edges \gets \left\{similarity(f_i, f_j) \mid getImg(f_i)
        \neq getImg(f_j) \wedge f_i, f_j \in p \right\}$
    \If{$\left\vert edges \right\vert > 1$}
        \State $matches \gets MMGet(p, t)$
    \ElsIf{$\left\vert edges \right\vert = 1$}
        \State $matches \gets RatioGet(p, features, t)$
    \Else
        \State $matches \gets \varnothing$
    \EndIf

    \Return matches
    \end{algorithmic}
\end{algorithm}

The matching algorithm for \emph{MMC} detailed in 
Algorithm~\ref{alg-getmatches} finds matches within all partitions with 
more than two elements using the \emph{MM} approach.  However, as can be 
seen in the example in Figure~\ref{fig:graph}, many of the partitions 
contain only two feature points from different images linked by one 
edge. In this case, we compare the similarity of the these two feature 
points with their second best match and remove matches where this ratio 
lies above a certain threshold, like in the \emph{Ratio} algorithm.  For 
example in the case of Figure~\ref{fig:pitts_partition}, we have feature 
points from a building in one image grouped together with points from a 
shirt pattern in another.  The nearest neighbor method would have 
returned wrong matches, but since we match the partition with \emph{MM}, 
points in the building end up matching other points in the building, and 
no matches are returned.

\section{The Competition}

The algorithms used in the comparison consists of two algorithms that 
use geometric data to enhance the matching, \emph{Isomatch} 
\cite{das2008event} and \emph{Spectral} \cite{leordeanu2005spectral} and 
one non geometric algorithm \emph{Ratio} \cite{lowe2004sift}.  
\emph{Ratio} is selected because it is a de facto standard for matching 
feature points, but also because it performs very well as shown by 
Mikolajczyk and Schmid in \cite{mikolajczyk2005performance}.  
\emph{Isomatch} and \emph{Spectral} on the other hand are selected 
because they are examples of two different approaches to geometric 
matching. \emph{Spectral} uses pairwise optimization to find a set of 
matches that is consistent in terms of direction and length which makes 
it less ideal in case of rotation and perspective change. 
\emph{Isomatch} focuses on matching only between geometric regions found 
by clustering the keypoints in the image.  This approach makes the 
algorithm resilient to rotation and perspective change, but will cause 
problems when the images matched are only partially overlapping and 
several regions do not have any corresponding regions in the other 
image.

\subsection{Ratio}

\begin{algorithm}[htb]
\caption{Ratio Match (\emph{Ratio})}
\label{alg-ratio}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $im_1$ : image, $im_2$ : image, $t \in \mathbb{R}$
\State $matches\gets \varnothing$
\State $features_1 \gets getFeatures(im_1)$
\State $features_2 \gets getFeatures(im_2)$
\ForAll{$f_i \in features_1$}
    \State $f_m,f_n \gets get2NNs(f_i, features_2)$
	\State $ratio \gets similarity(f_i, f_n) / similarity(f_i, f_m)$
	\If{$ratio < t$}
        \State $matches \gets \left(f_i, f_m\right)$
	\EndIf
\EndFor
\Return $matches$
\end{algorithmic}
%}
\end{algorithm}

\emph{Ratio} (Algorithm~\ref{alg-ratio}) \cite{lowe2004sift} is the most 
straightforward of the three reference algorithms. Feature points are 
matched using K Nearest Neighbor and the similarity ratio between the 
best and second best match is then compared to a threshold to decide if 
the match is kept.

\subsection{Isomatch}

\begin{algorithm}[h]
\caption{Isomatch (\emph{Isomatch})}
\label{alg-isomatch}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $im_1$ : image, $im_2$ : image, $t \in \mathbb{R}$
\State $matches_{init}\gets \varnothing$
\State $matches_{final}\gets \varnothing$
\State $features_1 \gets getFeatures(im_1)$
\State $features_2 \gets getFeatures(im_2)$
\ForAll{$f_i \in features_1$}
    \State $f_m,f_n \gets get2NNs(f_i, features_2)$
    \State $matches_{init} \gets \left(f_i, f_m\right)$
\EndFor
\State $partitions_1 \gets isodata(getPositions(features_1))$
\State $partitions_2 \gets isodata(getPositions(features_2))$
\State $P_{corr} \gets getMatchMat(partitions_1, partitions_2, 
matches_{init})$
\ForAll{$(i,j) \in indices(P)$}
    \If{$\left(P_{i,j} > 5\right) \wedge \left(P_{i,j} / sum(P_{i}) > 
    0.5\right)$}
        \State $matches_{final} \gets getMatches(matches_{init}, 
        partitions_1, partitions_2, i, j)$
    \EndIf
\EndFor
\Return $matches_{final}$
\end{algorithmic}
%}
\end{algorithm}

\begin{figure}[htb]
	\centering
	\includegraphics[width=\columnwidth]{images/isomatch_partitions}
	\caption{The result of partitioning feature points in two images.  
	Each dot represents a feature point and feature points grouped with 
the same color belong to the same partition.}
	\label{fig:isomatch_partitions}
\end{figure}

The \emph{Isomatch} algorithm (Algorithm~\ref{alg-isomatch}) 
\cite{das2008event} is based on the idea that local feature points 
within a region in one image often matches to features within a 
corresponding region in the other image. To find these regions the 
algorithm uses the Isodata algorithm \cite{ball1965isodata} to perform 
unsupervised clustering on the feature points given their positions in 
the two images. The Isodata algorithm is similar to K-means with the 
addition that clusters are merged and split during the clustering 
process. In Figure~\ref{fig:isomatch_partitions} an example of a 
partitioning can be seen. The initial set of matches are filtered based 
on the resulting regions. A match is kept only if it at least half the 
matches of the same region of origin also belong to the same region in 
the other image.  $P_{corr}$ in Algorithm~\ref{alg-isomatch} is a matrix 
where each row corresponds to a region in $im_1$, each column 
corresponds to a region in $im_2$ and $P_{i,j}$ is the amount of matches 
going from region $i$ to region $j$.

To threshold the \emph{Isomatch} algorithm we filter the results using 
the ratio between first and second best match as in \emph{Ratio}, since 
\emph{Isomatch} itself doesn't readily admit any thresholding parameter.

\subsection{Spectral}

\begin{algorithm}[h]
\caption{Spectral Match (\emph{Spectral})}
\label{alg-spectral}
%{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $im_1$ : image, $im_2$ : image, $t \in \mathbb{R}$
\State $matches_{init}\gets \varnothing$
\State $matches_{final} \gets \varnothing$
\State $features_1 \gets getFeatures(im_1)$
\State $features_2 \gets getFeatures(im_2)$
\ForAll{$f_i \in features_1$}
	\State $f_m,f_n \gets get2NNs(f_i, features_2)$
	\State $matches_{init} \gets \left(f_i, f_m\right)$
\EndFor
\State $M \gets matrix(\left\vert matches_{init} \right\vert, \left\vert 
matches_{init} \right\vert)$
\ForAll{$m_i \in matches_{init}$}
	\ForAll{$m_j \in matches_{init}$}
		\If{$i = j$}
			\State $M \gets matchSimilarity(m_i)$
		\Else
			\State $M \gets affinity(m_i, m_j)$
		\EndIf
	\EndFor
\EndFor
\State $x^{*} \gets maximumEigenvector(M)$
\ForAll{$\left(m_i, x_i\right) \in \left(matches_{init}, x^{*}\right)$}
	\If{$x_i > t$}
		\State $matches_{final} \gets matchtes_{final} \cup m_i$
	\EndIf
\EndFor

\Return $matches_{final}$
\end{algorithmic}
%}
\end{algorithm}

The \emph{Spectral} algorithm (Algorithm~\ref{alg-spectral}) is a 
slightly modified version of the spectral matching algorithm proposed by
Leordeanu and Herbert \cite{leordeanu2005spectral}. It is based on the 
idea that the set of actual correspondences of feature points between 
two images should be pairwise consistent. That is, given two feature 
points that are geometrically close in one image, for most cases we 
would assume that the closest correspondences in another image would be 
similarly close. We want to pick out a set of matches $T \subset S$ 
where $S$ is the initial set of possible matches such that the matches 
in $T$ are as pairwise consistent as possible. If we let $x$ be a binary 
vector corresponding in length to the set of initial matches $S$, where 
$x_i = 1$ if $m_i \subset T$, then we can formulate a subset-score as:
\begin{equation*}
	S = \sum_{m_i, m_j \in T} M(m_i, m_j) = x^TMx
\end{equation*}
The solution $x^{*}$ that maximizes this equation is then $x^{*} = 
argmax(x^TMx)$. By relaxing the constraints on $x$ such that it can take
values in the range of $\left[0, 1\right]$ we can interpret $x_i$ as the
association of $m_i$ with the optimal subset of matches $T$. Because we 
only worry about the relative values of $x$ we can fix the norm of $x$ 
to $1$, in which case by Raleigh's ratio theorem, the $x^{*}$ that 
optimizes $x^TMx$ is the principal eigenvector of $M$.

Leordeanu and Herbert proposes a simple affinity function to map the 
pairwise distance of two matches. In a scene where no rotation and 
excessive affine transformation takes place, it can generally be assumed 
that the length and direction of the vector between
the two feature points constituting the first match corresponds to the 
length and direction of the vector between the two feature points 
constituting the second match. Given two matches $m_i = (f_{i,1}, 
f_{i,2})$ and $m_j = (f_{j,1},f_{j,2})$ where $f_{k,n}$ is the feature 
point of match $k$ in image $n$, we denote the euclidean distance 
between feature points $f_{i,n}$ and $f_{j,n}$ in image $n$ as 
$d_{i,j,n}$.  The affinity function assigning a score based on the 
pairwise distance is then defined as follows:
\begin{equation*}
	M(a,b) \begin{cases} 4.5 - \frac{\left(d_{i,j,1} - 
		d_{i,j,2}\right)^2}{2\sigma^2}, & \mbox{if } \left\vert 
				d_{i,j,1} - d_{i,j,2} \right\vert < 3\sigma \\ 0, & 
				\mbox{otherwise}
	\end{cases}
\end{equation*}
In the equation above $\sigma$ is a stabilizing parameter that controls 
the sensitivity of the score to deformations in the data. A larger 
$\sigma$ will allow for larger deformations. In the experiments it is 
fixed to 50. There are a number of occasions where the assumption that 
pairwise matches should have the same distance doesn't hold true.  A 
simple example would be matches between two images of the same scene 
where one image is rotated $180^{\circ}$.  However for a lot of 
applications like panorama stitching where we only see minor rotation 
and change of perspective the assumption is accurate.

In the implementation proposed by Leordeanu and Herbert 
\cite{leordeanu2005spectral} the last step of the algorithm greedily 
selects matches to ensure that matches are consistent, i.e.\ that two 
feature points in one image don't match with the same feature point in 
the other image. This is done by going through all the matches in the 
order of their associated $x^{*}$ score and for each match eliminating 
all matches in conflict with this match. This method ensure that as many
matches are kept as possible since only the conflicting matches are 
removed. However for cases where most feature points in one image will 
not have a correspondence in the other other image (e.g.\ cases of 
partial image overlap or object recognition) the greedy matching 
algorithm will keep finding matches even after the actual 
correspondences have already been found as long as they are consistent.

Since the testing is done on images with partial overlap, the final set 
of matches returned in Algorithm~\ref{alg-spectral} is the set of 
matches where the association of the match with the optimal subset of 
matches $T$ as measured by $x^{*}$ is above a certain threshold. For 
experiments the set of initial matches were given by the best match per 
feature point and the threshold was adjusted to allow for the best 50\% 
to be included in the set of final matches. A rate less than 50\% 
usually did not improve results, so to threshold the \emph{Spectral}
algorithm we filter the output using the ratio between first and second
best match as done in \emph{Ratio}.