### 2013-08-13 text/head/abstracts.tex
A few touches on the thesis

%\begingroup
%\let\cleardoublepage\clearpage

% English abstract
\cleardoublepage
\chapter*{Abstract}
%\markboth{Abstract}{Abstract}
\addcontentsline{toc}{chapter}{Abstract} % adds an entry to the table of 
% put your text here
Many algorithms have been proposed to solve the problem of matching 
feature points in two or more images using geometric assumptions to 
increase the robustness of the matching. However, very little work 
addresses the case where these assumptions might not hold. In particular 
few methods address the problem of reliable matching in cases where it 
is unknown whether two images have any corresponding areas or objects in 
the first place.  \\

We propose two algorithms for matching feature points without the use of 
geometric constraints. The first relies on the fact that any match 
between two images should be better than all possible matches within a 
single image. The second algorithm extends this idea by using community 
structure in the similarity graph of feature points to find reliable 
correspondences. To evaluate the algorithms experimentally, we introduce 
a simple method to generate a large amount of test cases based on a set 
of image pairs with viewpoint changes. Our results show that the 
proposed algorithm is generally superior to traditional approaches in 
finding correct correspondences.