### 2013-08-13 text/head/titlepage.tex

\begin{titlepage}
\begin{center}
%\large
\sffamily

\null\vspace{2cm}
{\huge Reliable Feature Point Matching} \\[24pt]
{\small MASTER'S THESIS}

\vfill

\begin{tabular} {cc}
\parbox{0.3\textwidth}{\includegraphics[width=4cm]{images/adsc}\\[18pt]
\includegraphics[width=4cm]{images/epfl}}
&
\parbox{0.7\textwidth}{%
    Jonas Toft Arnfred\\
    Master's student in Computer Science\\
	at École Polytechnique Fédérale de Lausanne (EPFL)\\[9pt]
	Spring semester 2013\\
	Advanced Digital Sciences Center (Singapore)\\[9pt]
    Supervisors at ADSC: \ Dr. Stefan Winkler\\ [4pt]
    Supervisor at EPFL: ~ \ Prof. Sabine Süsstrunk\\ [4pt]
    \null \hspace{3em} }

%
%Singapore, ADSC, August 2013
\end{tabular}
\end{center}
\vspace{2cm}
\end{titlepage}