### 2014-04-15 text/journal/paper.tex
Stefans changes

%% bare_jrnl_compsoc.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE Computer
%% Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
% The Computer Society usually requires 10pt for submissions.
%
\documentclass[10pt,journal,cspaper,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[12pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress]{cite}
\else
  % normal IEEE
  \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
%\usepackage[tight,normalsize,sf,SF]{subfigure}
\else
%\usepackage[tight,footnotesize]{subfigure}
 \fi
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. Computer Society papers
% use a larger font and \sffamily font for their captions, hence the
% additional options needed under compsoc mode. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false]{caption}
%  \usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false]{caption}
%  \usepackage[font=footnotesize]{subfig}
%\fi
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

%\documentclass{article}
\usepackage{footnote}
\usepackage{multirow} 
%\usepackage[font={small}]{caption}
%\usepackage{sidecap}
\usepackage{cite}
\usepackage[draft]{hyperref}
%\usepackage[options]{nohyperref}
%\usepackage{url}
%\usepackage[stable]{footmisc}
%\usepackage{adjustbox}
\usepackage{balance}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
%\usepackage{authblk}
\pdfminorversion=4
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\newcommand{\twopartdef}[4]
{
	\left\{
		\begin{array}{ll}
			#1 & \mbox{if } #2 \\
			#3 & \mbox{if } #4
		\end{array}
	\right.
}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Decade of Lost Precision:\\
A General Framework For Image Feature Matching Without Geometric Constraints}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Jonas~Toft~Arnfred,
        Stefan~Winkler,
        Sabine~S\"usstrunk% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem J.~T.~Arnfred and S.~Winkler are with the Advanced Digital Sciences Center (ADSC), University of Illinois at Urbana-Champaign (UIUC), Singapore.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
% E-mail: 
\IEEEcompsocthanksitem S.~S\"usstrunk is with the \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland.}% <-this % stops a space
\thanks{}}

%\IEEEauthorblockN{%
%Jonas Toft Arnfred,\IEEEauthorrefmark{1} Stefan Winkler,\IEEEauthorrefmark{1} Sabine S\"usstrunk\IEEEauthorrefmark{2}}\\
%\vspace{1mm}
%\IEEEauthorblockA{%
%\IEEEauthorrefmark{1}~Advanced Digital Sciences Center (ADSC), University of Illinois at Urbana-Champaign (UIUC), Singapore\\
%\IEEEauthorrefmark{2}~\'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL), Switzerland}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
%\markboth{IEEE Trans. PAMI}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
Computer vision applications that involve the matching of local image 
features frequently use \emph{Ratio-Match} as introduced by Lowe and 
others, but is this really the optimal approach?  We formalize the 
theoretical foundation of \emph{Ratio-Match} and propose a general 
framework encompassing \emph{Ratio-Match} and five other matching 
methods. Using this framework, we establish a theoretical 
performance ranking in terms of precision and recall, 
proving that all five methods consistently outperform or equal \emph{Ratio-Match}.

We confirm the theoretical results experimentally on over 3000 image 
pairs and show that we can increase matching precision
without further assumptions about the images we are using.  These gains 
are achieved by making only a few key changes of the \emph{Ratio-Match} algorithm and do not increase computation times.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.

% Note that keywords are not normally used for peer review papers.
\begin{keywords}
Feature matching, ratio match, mirror match, self match
\end{keywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode
% is not selected <OR> if conference mode is selected - because compsoc
% conference papers position the abstract like regular (non-compsoc)
% papers do!
\IEEEdisplaynotcompsoctitleabstractindextext
% \IEEEdisplaynotcompsoctitleabstractindextext has no effect when using
% compsoc under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
%
Matching image points is a crucial ingredient in almost all computer 
vision algorithms that deal with sparse local image features. This 
includes image categorization \cite{bosch2008scene}, image stitching 
\cite{brown2007automatic}, object detection \cite{zhang2007local}, and 
near duplicate detection \cite{zhao2009scale}, to mention just a few examples.  
These problems all rely on being able to accurately find the 
correspondence(s) of a point on an object in a \emph{query image} given 
one or more \emph{target images} that might contain the same object.  In 
many applications the target images have undergone transformations with 
respect to the query image; in stereo vision, the viewpoint is 
different, while in object recognition and near duplicate detection both 
the lighting and even the object itself may also be transformed.

In the literature two approaches to feature point matching have been 
pursued and later merged. The geometric approach tries to find unique 
keypoints in an image pair and match them based on their location.  The 
descriptor centric approach on the other hand uses the local image 
information around the keypoint to create a descriptor and matches 
descriptors based on their similarity.

In the purely geometric approach, feature points are matched based on their position in the images. 
Scott and Longuet-Higgins 
\cite{scott1991algorithm} and Shapiro and Brady 
\cite{shapiro1992feature} introduced the use of spectral methods by 
deriving a coherent set of matches from the eigenvalues of the 
correspondence matrix. Other examples of this approach include \cite{sclaroff1995modal,carcassoni2003spectral}.

The descriptor centric approach on the other hand finds matches by 
pairing similar keypoints. The first 
examples of this approach used the correlation of the raw image data 
immediately surrounding the feature point \cite{deriche1994robust,baumberg2000reliable} to calculate this similarity. Later 
algorithms were enhanced by invariant feature descriptors as first 
introduced in \cite{schmid1997local} and later popularized by the work 
of Lowe introducing SIFT \cite{lowe2004sift} and Bay et al.\ introducing SURF 
\cite{bay2006surf}.

Many descriptor variations have since been proposed, and a few surveys 
have emerged comparing them. Mikolajczyk and Schmid \cite{mikolajczyk2005performance} evaluate 11 such 
feature descriptors in a paper where they also introduce the graffiti 
image set which has later become a common benchmark for image matching 
algorithms.  Moreels and Perona \cite{moreels2007evaluation}
evaluate 5 descriptors with different keypoint detectors using a large 
image set consisting of 3D models. 
Heinly et al.\ \cite{heinly2012comparative} take a closer look at binary features such as BRIEF 
\cite{calonder2010brief}, BRISK \cite{leutenegger2011brisk}, and ORB 
\cite{rublee2011orb}, comparing them to SIFT and SURF.

A straightforward way to find a set of correspondences using only 
feature points is to apply a threshold to the similarity measure of the feature 
vectors, accepting only correspondences that score above a certain 
level of similarity \cite{szeliski2010}. When we match images with the assumption 
that the correspondence between two feature points will be unique, we can 
further increase precision by only matching a feature point to its 
nearest neighbor in terms of descriptor similarity. Instead of 
thresholding based on similarity, Deriche et 
al.~\cite{deriche1994robust} and Baumberg~\cite{baumberg2000reliable} 
propose using the ratio of the similarity of the best to second best 
correspondence of a given point to evaluate how unique it is. Their 
finding has later been tested by several independent teams, all 
concluding that thresholding based on this ratio is generally superior to 
thresholding based on similarity \cite{lowe2004sift,mikolajczyk2005performance,moreels2007evaluation,rabin2009statistical}.

Brown and Lowe \cite{brown2005multi} extend ratio match to deal 
with a set of images by using not the ratio of the best and second best 
correspondence, but the average ratio of the best and the average of 
second best correspondences across a set of images.  Rabin et al.\ tries to enhance descriptor 
matching by looking at the statistical distribution of local features in 
the matched images, and only return a match when such a 
correspondence would not occur by mere chance 
\cite{rabin2009statistical}. Finally, a precursor of the algorithms 
discussed in this paper was introduced by the authors as 
\emph{Mirror-Match} \cite{arnfred2013mirror}, which makes use of the 
feature points in both images to decide if a match is valid.

A plethora of solutions have combined ratio match with various geometric 
constraints to improve matching. These constraints are
based on assumptions regarding the transformation between the query and 
target images. At the stricter end we have epipolar constraints, assuming
that the images can be tied by a homography \cite{torr2000mlesac,chum2005matching}, and angular constraints, assuming the 
correspondences are angled similarly \cite{kim2008efficient,schmid1997local,leordeanu2005spectral}. 
These constraints can be relaxed using more sophisticated models.  Torresani et al.\ \cite{torresani2008feature} 
attempted to model transformations using graph matching similar to 
\cite{deriche1994robust}. Others define image regions and reject or 
accept correspondences based on the regions they connect 
\cite{cho2009feature,wu2011robust}.

Any matching method relying on geometric constraints is restrained by 
inherent assumptions about the geometric relationship between the two 
images. Broad assumptions such as the epipolar constraint only apply in 
simple image transformations. For more complex transformations we need 
models suitable for each particular case, which limits the algorithm 
to the subset of images that fit the model.  In the case of object 
recognition for example, the transformations from one scene to another 
often feature a change in perspective, background, and sometimes variations within the object 
itself: a person can change pose, a car model can have different 
configurations, a flower can bloom etc.

When matching these instances we are forced to either create a 
sophisticated model that represents the variables of transformation 
within the object, or alternatively find correspondences using an 
algorithm with no inherent geometric assumptions. Besides, 
any geometric method acts as a filter on a given set of 
correspondences.  Therefore, if the initial set of purely descriptor-based matches contains fewer incorrect correspondences, 
the final set can be calculated faster and more accurately.

The methods we propose in this paper are designed to be free from 
assumptions about image geometry. They extend and improve on Lowe's 
\emph{Ratio-Match} \cite{lowe2004sift} and the authors' \emph{Mirror 
Match} \cite{arnfred2013mirror} by generalizing both algorithms to a 
framework of matching methods. We go on to formally establish a ranking
based on how well different methods within the framework compare in 
terms of precision and recall. Our experimental evaluations confirm the 
theoretical results and show that \emph{Ratio-Match} on all counts is generally a 
sub-optimal choice as a matching algorithm.

The paper is organized as follows. In Section~\ref{S:MatchingMethods} we 
introduce the original \emph{Ratio-Match}, generalize this method to 
introduce the proposed framework, and compare the methods of the 
framework theoretically.  In Section~\ref{S:Experiments} we present 
our experimental validation and discuss the results obtained.  Section~\ref{S:Summary} 
concludes the paper.

\section{Matching Framework}
\label{S:MatchingMethods}
%
\subsection{Ratio-Match}
%
\emph{Ratio-Match} as introduced by Deriche et al.\ 
\cite{deriche1994robust} and later used by 
Baumberg~\cite{baumberg2000reliable} and Lowe~\cite{lowe2004sift} makes 
use of a \emph{uniqueness ratio}. For each feature in the \emph{query 
image}, it finds a possible match by identifying the nearest neighbor and 
returning this match only if the \emph{uniqueness ratio} is above a 
given threshold $\tau$.

Given $f_q$, a feature from the \emph{query 
image} and $F_{target}$, a set of features from the \emph{target image}, 
the nearest neighbor of $f_q$ in $F_{target}$ is:
\begin{equation*}
    f_{nn} = \argmin_{f_t \in F_{target}} d(f_q, f_t).
\end{equation*}
Here, $d(f_i, f_j)$ is the distance between feature descriptors.  
With SIFT and SURF this is the Euclidean distance of the feature vectors
\cite{lowe2004sift,bay2006surf}, whereas BRIEF, BRISK, and FREAK 
use the Hamming distance \cite{leutenegger2011brisk,calonder2010brief,alahi2012freak}.

If we let $f_p$ and $f_b$ be the nearest and second nearest neighbor of a feature $f_q$, 
the \emph{uniqueness ratio} is defined as follows:
\begin{equation*}
    r = d(f_q, f_p) / d(f_q, f_b).
\end{equation*}
Because $f_p$ is the nearest neighbor and $f_b$ the second nearest 
neighbor, we know that $d(f_q, f_p) \leq d(f_q, f_b)$, and since $d(f_i, 
f_j) \geq 0, \forall f_i,f_j$, the \emph{uniqueness ratio} is bounded by $0$ 
and $1$, both inclusive. For the degenerate case where $d(f_q, f_b) = 0$
we let the \emph{uniqueness ratio} be $0$.

The underlying assumption in ratio match is that the point we seek to 
match in the \emph{query image} has only one match in a given 
\emph{target image} or no matches at all. In both cases we should expect 
the \emph{second} nearest neighbor in the target image to be an 
untrue correspondence.  This means we can use the distance between the 
feature point and the second closest descriptor as a \emph{baseline} for 
false correspondences. By dividing the distance to the nearest neighbor 
with this baseline, we get an estimated measure for how distinct the 
correspondence is from a false match.

To make further discussion easier we introduce the following 
nomenclature:
\begin{itemize}
\item{Let $f_q$ be any feature point in the \emph{query image} that we 
    are currently trying to match}.
\item{Let $F$ be a set of features. $F_{target}$ for example is the set 
    of features from the \emph{target image}}.
\item{Let $\tau \in [0 \ldots 1]$ be a threshold used to decide whether 
	to keep a match or discard it. We keep a match only if it has a 
uniqueness ratio lower than $\tau$}.
\item{Let the \emph{proposed match} be the nearest neighbor of $f_q$
    picked from a set of feature points that we will call the 
\emph{proposal set} which does not include $f_q$}.
\item{Let the \emph{baseline match} be the nearest neighbor of $f_q$ 
    picked from a set of feature points that we will call the 
\emph{baseline set}, which neither contains the \emph{proposed match} nor 
$f_q$}.
\end{itemize}

In the case of \emph{Ratio-Match} \cite{lowe2004sift}, the \emph{proposal 
set} consists of all the feature points in the target image.  Similarly 
the \emph{baseline set} consists of all the feature points in the target 
image except the \emph{proposed} match. For \emph{Ratio-Match} this 
practically means that given a feature $f_q$, the \emph{proposed match} 
and the \emph{baseline match} are the two nearest neighbors in the set 
of feature points in the \emph{target image}.

Brown et al.\ \cite{brown2005multi} extend \emph{Ratio-Match} for the case of $n$ target 
images by letting the baseline be the average of the distance to the 
second nearest neighbors for all $n$ target images.  They write, ``\emph{[W]e can improve outlier rejection by using information from all of the images (rather 
than just the two being matched). This improves the matching precision}''.  The idea of improving matching precision by expanding the 
knowledge used to reject outliers is also the central tenet of the framework presented in this paper.

%
\subsection{Framework of Matching Methods}
%

\emph{Ratio-Match} is only one instance of a framework of six algorithms we can 
construct by different combinations of \emph{proposed} and 
\emph{baseline sets}. As shown in Table~\ref{table:distinctness} for the 
case of two images, the \emph{proposal set} can be constructed by either 
looking only at the nearest neighbors in the target image, or 
alternatively by looking at the nearest neighbors in both the query and 
target image. The \emph{baseline set} can either be the query image 
itself, the target image, or the combination of the two.  Seen in this 
light, the matching algorithm \emph{Ratio-Match} draws both its 
\emph{proposal set} and \emph{baseline set} from the target image. In 
turn \emph{Mirror-Match-Ext} as introduced by the authors in 
\cite{arnfred2013mirror}, works by constructing the \emph{proposal set} 
and \emph{baseline set} by combining features from both the \emph{query} 
and \emph{target image}. Alternatively \emph{Self-Match} looks only to 
the query image for the \emph{baseline set} but constructs the 
\emph{proposal set} from the target image like \emph{Ratio-Match}. In 
Figure~\ref{fig:overview} we illustrate how each method is derived.

\begin{table}[htb]
\caption{Classification of the methods of the proposed framework based on the feature sources for 
    \emph{proposal set} and \emph{baseline set} for calculating the \emph{uniqueness ratio}.}
\label{table:distinctness}
	\centering
\bgroup
\def\arraystretch{1.5}
    \begin{tabular}{l l | l l}
        %
        %&  \multicolumn{1}{l}{Target}
        %& \multicolumn{1}{l}{Target and Query} \\
\multicolumn{2}{r}{} & \multicolumn{2}{c}{\emph{Proposal Set}} \\
                     & & Target & Target and Query \\
        \cline{2-4}
        \multirow{3}{*}{\rotatebox[origin=c]{90}{\emph{Baseline Set}}} &
  Query            & \emph{Self-Match}   & \emph{Self-Match-Ext} \\
& Target           & \emph{Ratio-Match}   & \emph{Ratio-Match-Ext} \\
& Target and Query & \emph{Mirror-Match} & \emph{Mirror-Match-Ext} \\
        \cline{2-4}
\end{tabular}
\egroup
\end{table}

\begin{figure}[htb]
\centering
\includegraphics[width=0.85\columnwidth]{images/overview}
\caption{Graphical representation of the \emph{baseline} and 
\emph{proposal sets} for the methods in the proposed framework.}
\label{fig:overview}
\end{figure}

Algorithm~\ref{alg-gen} shows the generalized matching framework, 
where the \emph{proposal set} and \emph{baseline set} are defined 
according to Table~\ref{table:distinctness}.  In practice for e.g.\ 
\emph{Ratio-Match} we can obtain $f_p$ and $f_b$ by finding the two 
nearest neighbors of $F_{target}$. The other algorithms can be optimized 
similarly. Figure~\ref{fig:matching} illustrates the structure of 
the generalized algorithm.

\begin{algorithm}[htb]
\caption{Generalized matching algorithm for two images}
\label{alg-gen}
\begin{algorithmic}
    \Require $I_{query}, I_{target}$ : images, $\tau \in [0,1]$
    \State $F_{query} = get\_features(I_{query})$
    \State $F_{target} = get\_features(I_{target})$
\State $F_{proposal} = get\_proposal\_set(F_{query}, F_{target})$
\State $F_{baseline} = get\_baseline\_set(F_{query}, F_{target})$
\State $M = \varnothing$
\ForAll{$f_q \in F_{query}$}
    \State $f_p \gets getNearestNeighbo\text{r}(f_q, F_{proposal} \setminus 
    \left\{f_q\right\})$
    \State $f_b \gets getNearestNeighbo\text{r}(f_q, F_{baseline} \setminus 
    \left\{f_q, f_p\right\})$
    \State $ratio \gets distance(f_q, f_p) / distance(f_q, f_b)$
    \If{$(ratio < \tau) \wedge (f_p \in F_{target})$}
        \State $matches \gets matches \cup \left(f_q, f_p\right)$
	\EndIf
\EndFor
\Return $M$
\end{algorithmic}
\end{algorithm}

\begin{figure}[htb]
\centering
\includegraphics[width=0.9\columnwidth]{images/matching}
\caption{Feature matching flow chart. $\tau$ 
is the ratio threshold, and $d(x,y)$ is the distance between two feature 
descriptors $x$ and $y$.}
\label{fig:matching}
\end{figure}
%

For some methods it is possible that the \emph{proposal match} is a 
feature from the \emph{query image}, in which case we discard the match 
as an untrue correspondence.  It also happens that we encounter 
correspondences with a \emph{uniqueness ratio} above $1$, in which case we set 
it to $1$ to maintain a bounded \emph{uniqueness ratio}.  Take for example the 
case of \emph{Self-Match}. Using this method we might find that the 
nearest neighbor of a feature in the target image is further from the 
query feature than the nearest neighbor in the query image. In this case 
the match is discarded.

\subsection{Generalized Uniqueness Ratio}

Form $F_{query} = \left\{f_{q_1} \ldots f_{q_n}\right\}$ and $F_{target} 
= \left\{f_{t_1} \ldots f_{t_m}\right\}$.  We can obtain $F_{proposed}$ 
and $F_{baseline}$ from these sets depending on the algorithm used (see 
Table~\ref{table:distinctness}). $r = \text{r}(f_{q}, F_{proposed}, 
F_{baseline})$ is the ratio of query feature $f_{q}$ given 
$F_{proposed}$ and $F_{baseline}$.  If $f_{p}$ is the nearest neighbor 
of $f_{q}$ in $F_{proposed}$ and $f_{b}$ is the nearest neighbor of 
$f_{q}$ in $F_{baseline} \setminus \left\{f_{p}\right\}$, then the 
generalized \emph{uniqueness} ratio is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \frac{d(f_{q}, f_{p})}{d(f_{q}, f_{b})}.
\end{align*}

We let $K$ be the number of possible true correspondences between \emph{query} 
and \emph{target image}. \emph{Precision} and \emph{recall} are 
defined as:
\begin{align*}
    \textrm{Recall} &= \frac{\#\textrm{Correct Matches}}{K} \\
    \textrm{Precision} &= \frac{\#\textrm{Correct 
    Matches}}{\#\textrm{Correct Matches} + \#\textrm{Incorrect Matches}}
\end{align*}
Given $f_{nn} \in F_{target}$ as the nearest neighbor of any 
$f_q \in F_{query}$, we define the set of features $F_{true}$ as 
$\left\{ f_{q} \in F_{query} \mid f_{nn} \text{ is a true correspondence 
of } f_{q} \right\}$ and $F_{false}$ as $F_{query} \setminus F_{true}$.  
We can then define the number of correct and incorrect matches as 
follows:
\begin{align*}
    \textrm{keep\_match}(f_{q}) &= \twopartdef{ 1 }{\text{r}(f_{q}, 
    F_{proposed}, F_{baseline}) <
    \tau}{0}{\textrm{otherwise}} \\
    \#\textrm{Correct Matches} &= \sum_{f_{q} \in F_{true}} 
    \textrm{keep\_match}(f_{q})\\
    %\mathbbm{1}_{\left\{ \right\}} \\
    \#\textrm{Incorrect Matches} &= \sum_{f_{q} \in F_{false}}
    \textrm{keep\_match}(f_{q})
    %\mathbbm{1}_{\left\{ r < \tau\right\}}
\end{align*}

From the above we can conclude that the performance in terms of 
precision and recall of any method from the proposed framework is 
uniquely identified by the \emph{uniqueness ratio} $r$.  If for every query 
feature $f_{q}$, $r$ is identical across two matching methods, then they 
return identical results.

\subsection{Proofs of Algorithm Performance}
\label{S:Proofs}

Under the following three assumptions, which largely reflect the performance of 
matching feature points in practice, we can theoretically compare the 
performance of the different algorithms shown in 
Table~\ref{table:distinctness}:
\begin{enumerate}
    \item{For a query and target image we assume that for any point in 
        the query image there is at most one real correspondence in the 
    target image and no real correspondences in the query image itself 
as assumed by \emph{Ratio-Match}.}
    \item{We assume that distances of feature descriptors are well 
            behaved.  More precisely, given $f_q$, a feature from the 
            \emph{query image}, and $f_{match}$, the real correspondence 
            from the set of features in the \emph{target image} 
            $F_{target}$, we have $\forall f_i \in F_{proposal}, 
            f_i \neq f_{match}: d(f_q,f_{match}) < d(f_q, f_i)$.}
    %\item{We assume that features within one image resemble each other 
            %more than features in different images. That is, given 
            %$f_a,
    %        f_i \in F_1$ and $f_b, f_j \in F_2$ where $(f_i, f_j)$ and 
    %    $(f_a, f_b)$ are true correspondences, and $F_1$ and $F_2$ are 
    %features collected from different images then we expect that 
            %$d(f_a, f_i)% < d(f_a, f_j)$ and $d(f_b, f_j) < d(f_b, 
            %f_i)$}
     \item{For a set of query features $F_{query}$ with a true 
             correspondence in the \emph{target image}, we assume that 
             the distribution of \emph{uniqueness ratios} using $F_{target}$ as 
             the baseline set is similar to using $F_{query}$, since the 
         two images in the case of true correspondences are guaranteed 
     to share part of the same scene.}
    \end{enumerate}

Based on these assumptions we first prove that the results of 
\emph{Self-Match} are identical to \emph{Self-Match-Ext}. To do this, we 
look at the nearest neighbor $f_{nn}$ of every query feature $f_{q}$ and 
consider the two cases of $f_{nn} \in F_{query}$ and $f_{nn} \in
F_{target}$. For $f_{nn} \in F_{target}$ the \emph{uniqueness ratio} of 
\emph{Self-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{query} \cup F_{target}, F_{baseline})\\
        &= \text{r}(f_{q}, F_{target}, F_{baseline}).
\end{align*}
Since $F_{proposed} = F_{target}$ for \emph{Self-Match}, this shows that
when the nearest neighbor is found in the target image, the two 
algorithms behave identically. For $f_{nn} \in F_{query}$ 
\emph{Self-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{target}, F_{query}) \\
        &\geq 1.
\end{align*}
The closest neighbor is found in the query image, so $d(f_{q}, f_{nn}) < 
d(f_{q},f_{p})$, where $f_{p}$ is the nearest neighbor from 
$F_{proposal}$. \emph{Self-Match-Ext} does not return a correspondence 
found in the \emph{query image}, and since $\tau \in [0,1]$, neither does 
\emph{Self-Match}. This proves that 
\emph{Self-Match} returns the same matches as \emph{Self-Match-Ext} for any set of query features. For 
this reason both algorithms will be referred to as just 
\emph{Self-Match}.

The above proof can easily be extended to show that \emph{Mirror-Match} and \emph{Mirror-Match-Ext} also return the same correspondences, and 
both will be referred to as \emph{Mirror Match} from here on.

For the case of \emph{Ratio-Match} and \emph{Ratio-Match-Ext}, we 
can show that \emph{Ratio-Match-Ext} is equal or better than 
\emph{Ratio-Match}.  Consider again the nearest neighbor $f_{nn}$ of a 
query feature $f_{q}$, and the two cases where we either have $f_{nn} 
\in F_{query}$ or $f_{nn} \in F_{target}$. For $f_{nn} \in F_{target}$ 
the \emph{uniqueness ratio} of \emph{Ratio-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{query} \cup F_{target}, F_{baseline})\\
        &= \text{r}(f_{q}, F_{target}, F_{baseline}).
\end{align*}
Since $F_{proposed} = F_{target}$ for \emph{Ratio-Match} this shows that
when the nearest neighbor is found in the target image, the two 
algorithms behave identically. For $f_{nn} \in F_{query}$ 
\emph{Ratio-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{target}, F_{target}).
\end{align*}
While \emph{Ratio-Match-Ext} does not return the correspondence if 
the nearest neighbor of a query feature is in the \emph{query image}, 
\emph{Ratio-Match} is likely to do so, since the \emph{uniqueness ratio} is 
found by dividing with the second nearest neighbor in the \emph{target 
image}. This shows that for cases where a query feature has a nearest 
neighbor in the \emph{query image}, \emph{Ratio-Match-Ext} is superior 
to \emph{Ratio-Match}. Since query features with a nearest neighbor in 
the \emph{query image} are false correspondences per assumptions \#1 and 
\#2, this proves that \emph{Ratio-Match-Ext} has superior 
\emph{Precision} to \emph{Ratio-Match} while maintaining equal 
\emph{Recall}.

Finally we show that the accuracy of \emph{Mirror-Match} is equal to or better than 
\emph{Ratio-Match-Ext}. Consider as before the nearest neighbor $f_{nn}$ 
of a query feature $f_{q}$. When $f_{nn}$ resides in the \emph{query 
image}, both algorithms behave alike and discard the match.  However, 
consider the \emph{uniqueness ratio} of \emph{Mirror Match} for the case where 
$f_{nn}$ resides in the \emph{target image}:
\begin{align*}
    r &= \text{r}(f_{q}, F_{proposed}, F_{baseline}) \\
        &= \text{r}(f_{q}, F_{proposed}, F_{query} \cup F_{target})\\
        &= max( \text{r}(f_{q}, F_{proposed}, F_{target}), 
    \text{r}(f_{q}, F_{proposed}, F_{query}) ).
\end{align*}
For a true correspondence the \emph{uniqueness ratio} using 
$F_{query}$ as a baseline is distributed similarly to the ratio when using 
$F_{target}$ according to assumption \#3. In this case the algorithm performs like 
\emph{Ratio-Match-Ext}.  However, for an untrue correspondence with a 
baseline match in the \emph{query image} which is closer than the
baseline match in the \emph{target image}, \emph{Mirror Match} will 
return a worse \emph{uniqueness ratio}. This in turns means that \emph{Mirror 
Match} is equal or better than \emph{Ratio-Match-Ext} in terms of  
\emph{Precision} while maintaining equal \emph{Recall}.

We have thus proven that -- based on assumptions \#1-3 -- for any given recall rate, the 
precision of the algorithms in the framework compares as follows:
\begin{align*}
    \textit{Self-Match} &= \textit{Self-Match-Ext} \\
    \textit{Mirror-Match} &= \textit{Mirror-Match-Ext} \\
    \textit{Ratio-Match} &\leq \textit{Ratio-Match-Ext} \leq 
    \textit{Mirror Match}
\end{align*}

The difference between \emph{Self-Match} and \emph{Ratio-Match} is  
the source of the \emph{baseline set}. For true correspondences 
they should perform similarly, otherwise the 
performance of the two depends entirely on the images being matched. 
When we match images where the target image may not overlap at all 
with the query image, using the query image and not the target image as 
a \emph{baseline set} seems like a reasonable choice given that the 
baseline set in this case would be closer to the feature matched and 
more strictly rule out untrue correspondences.

\subsection{Discussion of Assumptions}
\label{ref:disc_assumptions}

% Assumption 1
%For a query and target image we assume that for any point in the query 
%image there is at most one real correspondence in the target image and 
%no real correspondences in the query image itself as assumed by 
%\emph{Ratio-Match}

% Assumption 2
%We assume the distances of feature descriptors behave in a \emph{nice} 
%manner. That is, given $f_q$, a feature from the \emph{query image} and 
%$f_{match}$, the real correspondence from the set of features in the 
%\emph{target image} $F_{target}$, we have that $\forall f_i \in 
%F_{proposal}, f_i \neq f_{match}: d(f_q,f_{match}) < d(f_q, f_i)$

% Assumption 3
%For a query feature, $f_q$ with a true correspondence in the 
%\emph{target image}, we assume that the \emph{uniqueness ratio} using 
%$F_{target}$ is similar to the \emph{uniqueness ratio} using $F_{query}$, 
%since the two images in this case are guaranteed to share part of the 
%same scene

The three assumptions underlying the formal ranking of 
algorithms presented above have been chosen to reflect conditions under which we would ideally
match feature points. However, for each assumption there exist 
corner cases where it is no longer valid. In this section we 
discuss these corner cases in order to review the 
circumstances under which the absolute or relative performance of the 
algorithms might differ.

The first assumption states that there is at most one unique match to 
every feature point, which is often the case for natural images. 
However, for the recognition of object classes or in generated images 
with repetitive content, a point in the \emph{query image} might have 
several true correspondences in the \emph{target image}, so using a 
\emph{baseline match} from the target image will lead to a much higher 
\emph{uniqueness ratio}. However, as long as the query image does not 
contain repetitive objects, we can still use \emph{Self-Match} without 
loss of precision.

According to the second assumption, feature descriptors should behave such that
when matched, a true correspondence to a query feature will always be 
the nearest neighbor. For practical implementations of feature detectors and descriptors, 
this property strongly depends on the differences between \emph{query} 
and \emph{target image}, as shown in  
\cite{mikolajczyk2005performance,tuytelaars2008local}. In the degenerate 
case where the closest match is not an actual correspondence, there is 
no reason why this match should be disproportionately closer to the 
\emph{query feature} than any matches in the \emph{target image}, so in 
most cases we would expect the \emph{uniqueness ratio} to be high 
(signifying a bad match). This means that with a lenient threshold (close 
to 1), the guarantees about precision and recall equivalence between the 
different ratio algorithms might no longer hold due to the inconsistent descriptor behavior. 
It is worth noting that if 
assumption \#2 holds, matching using only nearest neighbor is not 
necessarily optimal, unless there is a complete scene overlap between
the images matched. For cases where there is little or no overlap, a 
nearest neighbor would indiscriminately return false matches even with 
well-behaved feature descriptors.

The third assumption states that the \emph{uniqueness ratios} using either the 
\emph{query} or \emph{target image} as the baseline set would be similar
in distribution given that the feature point we are matching has a true 
correspondence. We can support this assumption by looking at the 
\emph{uniqueness ratios} returned by \emph{Self-Match} and 
\emph{Ratio-Match}, since they use the \emph{query} and 
\emph{target image} respectively to calculate the \emph{uniqueness ratio}.  
Figure~\ref{fig:ratio_hist} shows the actual \emph{uniqueness ratios} 
measured from 3024 image pairs featuring 3D objects (see below for more on this dataset).  
For ratios lower than $0.7$, the \emph{uniqueness ratios} are similar, but as we approach 
more lenient thresholds, the ratios based on the \emph{query image} are 
higher than those for the \emph{target image}. This means that the third assumption 
is invalid for more lenient thresholds, and we can no longer expect \emph{Mirror-Match} to outperform 
\emph{Ratio-Match-Ext} in terms of precision without a penalty in recall.

\begin{figure}[htb]
\centering
\includegraphics[width=0.99\columnwidth]{images/results_ratio_hist}
\caption{Normalized histogram of the \emph{uniqueness ratios} of true 
correspondences for 84 different 3D objects.}
\label{fig:ratio_hist}
\end{figure}

\section{Experiments and Results}
\label{S:Experiments}
%
In this section we present the experiments and evaluation results comparing 
\emph{Self-Match}, \emph{Ratio-Match}, \emph{Ratio-Match-Ext} and 
\emph{Mirror-Match}. We use two different databases to compile 
results over different image variations.

To measure results on realistic and 
variate non-planar images, we test the algorithm on the 3D Objects 
database released by Moreels and Pietro \cite{moreels2007evaluation} in 
Section~\ref{S:3dobjects}. This database contains a set of 86 3D objects 
photographed from all sides at 5 degree intervals from two different 
elevation angles and under three different lighting conditions.
To measure the performance over variations such as viewpoint change, 
blur, compression noise, rotation, and zoom, we use the graffiti 
image set introduced by Mikolajczyk and Schmid 
\cite{mikolajczyk2005performance} in Section~\ref{S:Planar}.  This 
dataset contains 8 scenes each with 6 images linked by homographies and 
features only planar or almost planar scenes.

For the experiments we use the \emph{SIFT} descriptor and keypoint 
detector with default parameters as implemented in the OpenCV library 
2.4.6.  To further test how the framework performs across descriptors, we 
evaluate SURF \cite{bay2006surf}, BRISK \cite{leutenegger2011brisk}, 
BRIEF \cite{calonder2010brief}, and FREAK \cite{alahi2012freak} in 
Section~\ref{label:desc}.

Across all experiments, only the luminance channel is used in the feature detection and description stage.  
Images larger than $1024\!\times\!768$ are resized to fit within those dimensions using 
ImageMagick with default parameters.

\subsection{Experiments on 3D Objects}
\label{S:3dobjects}

\begin{figure*}[t]
	\centering
    \includegraphics[width=2\columnwidth]{images/results_all_objects}
    \caption{Results for the 3D objects dataset. Each plot 
    contains data accumulated from 84 objects photographed under 3 
different lighting conditions.}
    \label{fig:all_objects}
\end{figure*}

The 3D objects dataset by Moreels and Pietro \cite{moreels2007evaluation} allows us to experimentally compare matching 
algorithms over a large range of object and surface types rotated on a 
turnstile and photographed from every 5 degree turn.  15 objects from the dataset are shown in Figure~\ref{fig:3d_objects}.  We use images of 
84 different objects under three different lighting conditions at 12 
different angle intervals, conducting experiments with a total of 3024 
image pairs.

\begin{figure}[htb]
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/1}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/2}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/3}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/4}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/5}
    \end{subfigure}%
    \vspace{1.5 mm}

    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/6}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/7}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/8}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/9}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/10}
    \end{subfigure}%
    \vspace{1.5 mm}

    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/11}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/12}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/13}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/14}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.19\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/3d/15}
    \end{subfigure}%
    \vspace{1.5 mm}

    \caption{15 objects from the 3D Objects dataset by Moreels
    and Pietro \cite{moreels2007evaluation}.}
    \label{fig:3d_objects}
\end{figure}

To validate matches, Moreels and Pietro propose a method 
using epipolar constraints \cite[p.266]{moreels2007evaluation}, which is 
outlined in Figure~\ref{fig:frog}.  According to their 
experiments, these constraints are able to identify 
true correspondences with an error rate of $2\%$. We use their proposed 
method to generate the ground truth for the evaluation of our framework.

\begin{figure}[htb]
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_A}
        \caption{Query Image}
        \label{fig:frog_a}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad a 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_B}
        \caption{Auxiliary Image}
        \label{fig:frog_b}
    \end{subfigure}%
    \\ %add desired spacing between images, e. g. ~, \quad, \qquad 
    %a blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C}
        \caption{Target Image}
        \label{fig:frog_c}
    \end{subfigure}%
    \quad %add desired spacing between images, e. g. ~, \quad, \qquad 
    %blank line to force the subfigure onto a new line)
    \begin{subfigure}[t]{0.48\columnwidth}
        \centering
        \includegraphics[width=\columnwidth]{images/Frog_C_lines}
        \caption{Target Image with epipolar constraints}
        \label{fig:frog_c_lines}
    \end{subfigure}%
    \caption{Creating epipolar constraints based on three source images \cite{moreels2007evaluation}.  
        (a) \emph{Query image} marked with the position 
        of the feature we are attempting to match. (b) 
        Auxiliary image, taken at the 
        same rotation as the \emph{query image} but from a higher 
        viewpoint. Feature points near the epipolar line 
        of the feature point from the \emph{query image} are shown as dots.
				(c) Sample \emph{target image} with feature points near the epipolar line of the feature point from the \emph{query image}.
				(d) Same \emph{target image} with epipolar lines from all the 
    feature points shown in the \emph{auxiliary image}. 
		True correspondences should lie close to the intersections 
    of the epipolar line from (c) with the lines from (d). In this particular case two 
    feature points are possible correspondences.
}%
    \label{fig:frog}%
\end{figure}%

To compute the total number of possible correspondences, we take each 
feature in a \emph{query image} and count how many of them have a 
feature in the \emph{target image} which would satisfy the epipolar 
constraints outlined above. When using this dataset, features 
with no correspondences were not included in the set of features for testing, 
so as to avoid matching non-moving background and 
foreground objects.

We evaluate all matching algorithms from our framework on the 3D objects dataset by 
matching images at different angular intervals. For each object we 
pick the \emph{query} image as the image taken at 10 degrees rotation 
for calibration stability.  We then match this image with 
the same object turned an additional $\Delta$ degrees, $\Delta \in \{5, 10, 
\ldots, 60\}$.  For every angle interval we compare images taken under 
3 different lighting conditions as provided by the dataset.
We include all objects in the database for which photos at 5 degree 
angle intervals are available except for the ``Rooster'' and ``Sponge'' 
objects due to image irregularities. For each plot we show the 
accumulated results on all 3D objects, weighted by the number of possible 
true correspondences for the particular object. This ensures that each 
object contributes equally to the final result, despite some objects
resulting in disproportionally more matches than others.

Figure~\ref{fig:all_objects} shows the performance of the different matching methods in our proposed
framework for $12$ increasingly bigger angle differences. The results are shown 
in a precision / recall plot to make it easy to compare performance in 
terms of precision at similar levels of recall.  \emph{Ratio-Match} and \emph{Self-Match} show similar 
results, while \emph{Mirror-Match} and \emph{Ratio-Match-Ext} outperform 
both of them, showing the advantage of composing the \emph{proposed set} of 
features from both images. \emph{Mirror-Match} fares slightly but
consistently better than \emph{Ratio-Match-Ext}. In general we 
see the largest performance improvements at lower recall and a 
convergence of precision at higher recall. This is expected since a 
higher recall is a direct consequence of a more lenient threshold; as
we let the threshold approach $1$, we lose the benefits of thresholding 
on the \emph{uniqueness ratio}, and all methods start approximating the 
results of a simple nearest neighbor match. For all but 
\emph{Ratio-Match}, the features with better matches within the same 
image are still weeded out, which explains the tail end of 
\emph{Ratio-Match} at high recall where the other algorithms no longer 
have results. As we demonstrated in \cite{arnfred2013mirror}, the removal 
of within-image matches improves performance on image pairs with partial 
or no overlap. For no overlap we would expect a matching algorithm to 
reject matches within regions that have no matching counterparts in the 
other image.

The performance gap between methods increases gradually with angle, reaching its maximum between $25$ and $40$ degrees, 
where \emph{Mirror-Match} exhibits about 20 percentage-points higher precision than \emph{Ratio-Match}. 
At larger viewpoint differences above $45$ degrees, the performance gain decreases.  We suspect this is partly because assumption \#2 stops being valid when the images are transformed beyond 
a certain level of recognizability.

To summarize, the methods perform as predicted on the 3D Objects dataset
with \emph{Mirror-Match} in general performing better than 
\emph{Ratio-Match-Ext}, which in turn outperforms
\emph{Ratio-Match}. The experimental results also show that 
\emph{Self-Match} overall performs equally well as or better than 
\emph{Ratio-Match}.

\subsection{Planar Scenes}
\label{S:Planar}

\begin{figure*}[t!]
	\centering
    \includegraphics[width=2\columnwidth]{images/results_acr_all}
    \caption{Accumulated results for each scene in the image dataset by 
    Mikolajczyk and Schmid \cite{mikolajczyk2005performance}.}
    \label{fig:acr_objects}
\end{figure*}

To evaluate local feature descriptors, Mikolajczyk and Schmid \cite{mikolajczyk2005performance} introduced
a series of $8$ image sets featuring different transformations such as 
compression noise, rotation and zoom, blur and viewpoint change. The 8 source scenes are shown in Figure~\ref{fig:acr_images}.  
All  images are tied by homographies, which makes them convenient for testing 
purposes, but as a consequence they all depict planar scenes, which is 
somewhat limited and unrealistic. For each scene, there are 6 different 
images with increasing levels of a certain transformation, such that 
the change between image 1 and image 2 in a scene is only slight, while image 6 is markedly different from image 1 for most cases.

\begin{figure}[htb]
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/bark}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/bikes}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/boat}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/graf}
    \end{subfigure}%
    \vspace{1.5 mm}
    % Linebreak
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/leuven}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/trees}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/ubc}
    \end{subfigure}%
    ~ %
    \begin{subfigure}[t]{0.24\columnwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/acr/wall}
    \end{subfigure}%
    \caption{Scenes from the image dataset by Mikolajczyk and Schmid 
    \cite{mikolajczyk2005performance}, ordered as they appear in 
Figure~\ref{fig:acr_objects}.}
    \label{fig:acr_images}
\end{figure}

We match the first image in each scene with images 3, 4, 5, 
and 6 and show the accumulated result weighted by the possible number of 
true correspondences for each pair. We exclude the comparison with image 
2 because it is trivial in most cases and does not tell us much about 
the relative performance of the different methods. We find inliers by 
using the homography between image pairs as follows: Given a potential 
match between two pixels $p_1$ and $p_2$, $m = \left(p_1, p_2\right)$, 
and a homography $H$ relating the two images $I_1$ and $I_2$, we can 
calculate if $m$ is an inlier by checking if the two points satisfy the 
following criterion:
\begin{equation*}
\left\vert H p_1 - p_2 \right\vert + \left\vert H^{-1}p_2 - p_1 \right\vert < d_{\max},
\end{equation*}
i.e.\ the distance between $p_1$ translated to $I_2$ and $p_2$ 
\emph{plus} the distance between $p_2$ translated to $I_1$ should be 
less than a certain threshold (we use $d_{\max}=10$ pixels here).

Figure~\ref{fig:acr_objects} shows the results plotted with precision on 
the y-axis and recall along the x-axis. For scenes such as `bark', `boat' 
(zoom and rotation), `leuven' (lighting changes) and `wall' (viewpoint changes), all methods 
in the framework show similar results, largely because the precision approaches 100\% for low recall, which does not leave much room 
for improvement. For `bikes' and `trees' (blur) as well as `ubc'
(compression) and `graf' (viewpoint), \emph{Mirror-Match} clearly 
outperforms \emph{Ratio-Match}, albeit with a smaller margin than 
for the 3D objects dataset. Because many images in the dataset are 
trivial to match and only a few are very difficult, there is less difference
between the algorithms than with the 3D objects dataset. However, the 
theoretical predictions are still confirmed, with \emph{Mirror-Match} 
performing better than or equal to \emph{Ratio-Match-Ext}, which in turn 
performs better than or equal to \emph{Ratio-Match}. \emph{Self-Match} 
again achieves a performance comparable to \emph{Ratio-Match} using only the \emph{query image} as \emph{baseline 
set}.

\subsection{Descriptor Evaluation}
\label{label:desc}

\begin{figure*}[t!]
    \begin{subfigure}[t]{1\textwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/results_descriptors_imageset1}
		\caption{Keypoint / Descriptor combinations evaluated on 15 
		pairs of photos of 3D objects taken 25 degrees apart.}
        \label{fig:descriptors}
    \end{subfigure}%

    \begin{subfigure}[t]{1\textwidth}
        \centering
        \includegraphics[width=1\columnwidth]{images/results_descriptors_easy}
        \caption{Keypoint / Descriptor combinations evaluated on images 1 and 3 from the `Bikes' 
        scene.}
        \label{fig:descriptors_easy}
    \end{subfigure}%
    \caption{Evaluation of Keypoint / Descriptor combinations (K = Keypoint, D = 
Descriptor).}
\end{figure*}

To evaluate whether the improvement gains with the SIFT descriptor 
translate to other feature descriptors such as SURF \cite{bay2006surf} 
or binary features such as BRIEF \cite{calonder2010brief}, BRISK 
\cite{leutenegger2011brisk}, or FREAK \cite{alahi2012freak}, we evaluated 
the different descriptors on the 3D objects data set.  The binary 
features were each paired with the FAST feature detector for 
consistency, while SIFT and SURF were tested with their respective 
feature detectors.  For each keypoint detector and feature descriptor 
pair we used the standard implementation from OpenCV 2.4.6.

The descriptors were evaluated on 15 images of the 3D object dataset 
with a fixed angle offset of 25 degrees (pictured in 
Figure~\ref{fig:3d_objects}).  Otherwise the evaluation performed in the same manner as
the general evaluation of 3D Objects in Section~ref{S:3dobjects}, and the data presented is 
accumulated over the 15 objects, each set of results weighted by the 
possible number of correspondences.

The results in Figure~\ref{fig:descriptors} are again in line with the theoretical performance analysis of Section~\ref{S:Proofs} .
They also show a clear difference in the variance of matching method 
performance between SIFT and the rest of the descriptors. The 
performance of \emph{Mirror-Match} using SIFT shows a clear increase in 
precision over \emph{Ratio-Match} at equal recall levels. For other 
descriptors the improvement is much less pronounced.  For SURF, BRISK, 
and BRIEF we see a small improvement in precision using 
\emph{Mirror-Match} over \emph{Ratio-Match} and \emph{Ratio-Match-ext}, 
whereas FREAK yields similar results no matter what algorithm is used.  
For all descriptors \emph{Self-Match} performs on par with 
\emph{Ratio-Match}.

Part of this difference between SIFT and the 
other descriptors can be explained by looking at the recall rate 
with for example SURF\@. Given the same amount of possible matches, 
\emph{Ratio-Match} with SURF spans a larger range of recall.  
Since all other methods filter out matches that are better matched 
within the same image, but show no increase in precision, we can conclude 
that correct matches are discarded because there are better matches 
within the same image. This violates assumption \#2 stating that a 
descriptor is well behaved and matches best with a descriptor of a true 
correspondence. This means that for this particular evaluation case, SURF 
and the binary features tested are not sufficiently discriminative to 
provide a viable case for preferring \emph{Mirror-Match} over 
\emph{Ratio-Match}.

To test this hypothesis we looked at the results from matching an image 
pair with very little inter image variation. We compare images 1 and 
image 3 from the `bikes' scene in the image set released by Mikolajczyk 
and Schmid \cite{mikolajczyk2005performance}. The results 
in Figure~\ref{fig:descriptors_easy} show a market improvement for 
\emph{Self-Match}, \emph{Mirror-Match}, and \emph{Ratio-Match-Ext} over 
\emph{Ratio-Match} for the case of BRISK, BRIEF, and FREAK as expected.  
For SURF however, the algorithms perform similarly at low recall while 
\emph{Ratio-Match} shows better performance at more lenient thresholds.

\subsection{Performance}

In terms of computational complexity, the algorithms can be implemented 
in $O(n\log n)$, if we assume that both the \emph{query} and the 
\emph{target} image have $n$ feature points. For a target image with $m$ 
features where $m$ is significantly different from $n$, the complexities 
are as noted in Table~\ref{table:running_times}. 
In practice the constant factors 
involved in the actual matching of two images with approximately the 
same number of feature points are usually small enough that all the 
algorithms run at very similar speeds. This is also shown in 
Table~\ref{table:running_times}, which contains the running times as measured 
while matching 15 3D objects under three different lighting conditions 
($42$ image pairs were matched in total). The running times are averaged 
over three separate runs of each algorithm implemented using the same 
data structures and libraries in Python.

\begin{table}[htb]
\caption{Complexity and running times tested on 15 image pairs on 3 
lighting conditions with average $n = 236.7$ and average $m = 247.4$ feature points,
as tested on an Intel\textregistered\ Core\texttrademark\ i5-3550 CPU @ 
3.30~GHz with 8~GB memory.}
\label{table:running_times}
	\centering
%	\small
    \begin{tabular}{{l}{l}{l}}
    Algorithm & Complexity & Running Time\\
    \hline
    \noalign{\smallskip}
    %
    \emph{Self-Match} & $O(n\log(nm))$ & 118s  \\
    \emph{Self-Match-Ext} & $O(n\log(n(n+m)))$ & 109s\\
    \emph{Ratio-Match} & $O(n\log(m))$ & 114s\\
    \emph{Ratio-Match-Ext} & $O(n\log(n+m))$ & 112s\\
    \emph{Mirror-Match} & $O(n\log(n(n+m)))$ & 116s\\
    \emph{Mirror-Match-Ext} & $O(n\log(n+m))$ & 110s \\
    \hline
\end{tabular}
\end{table}

\subsection{Extending to Multiple Target Images}
%
When we match a query image against multiple target images, as 
for example in image stitching, assumption \#1 that any feature in the 
query image has at most one real correspondence in the set of target 
features is no longer true. It might be the case that two target images 
contain the same real object which yields two target features that are 
both real correspondences. In this case calculating the \emph{uniqueness 
ratio} by dividing the distance of one real correspondence with the 
distance to another renders it entirely useless if the two features are 
found in sets where there might be more than one correct 
correspondence.

Brown et al.\ \cite{brown2005multi} propose calculating the uniqueness 
ratio by taking the average of the $n$ nearest neighbors in each of the 
$n$ target images. Since their particular case deals with images that 
are aligned horizontally one after the other, this approach is not 
flawed per se. However for the more general case of $N$ target images 
with an unknown amount of overlap in the target group, their approach 
would not work since several features in the target group could possibly 
be a correct correspondence.

This leaves us with two alternatives. Either we match the query image to 
every target image individually, or we use a matching strategy where the 
\emph{baseline set} is guaranteed not to contain any real 
correspondences. \emph{Self-Match} provides such a set, since the 
\emph{baseline set} of \emph{Self-Match} only contains features in the 
query image itself. In addition, matching a query image to $N$ target 
images with \emph{Self-Match} is vastly faster than matching the query 
image to every target image: If we assume 
that every of the $N$ images has $n$ features, then matching the query 
image to all $N$ images can be done in $O(Nn\log(n))$ using a metric 
tree. However, matching with \emph{Self-Match} is done in $O(n\log(nN)) =
O(n\log(N) + n\log(n))$.

\section{Conclusion}
\label{S:Summary}

We have proposed a general framework of matching methods; building on the ideas 
behind \emph{Ratio-Match} and \emph{Mirror-Match}, we also introduced the 
variations \emph{Self-Match} and \emph{Ratio-Match-Ext}.

We formally proved under three assumptions that \emph{Mirror-Match} 
performs better than or equal to \emph{Ratio-Match-Ext} in terms of 
precision and recall, which in turn performs better than or equal to 
\emph{Ratio-Match}.

From an evaluation using images of rotated 3D objects, we have shown that
the theoretical conclusions are reflected in the experimental results with 
\emph{Mirror-Match} often outperforming \emph{Ratio-Match}
significantly over 3024 image pairs. We further confirmed this behavior 
for the classical `graffiti' image set, although there is less variation 
between the algorithms. Finally we show that the theoretical conclusions 
hold across a variety of popular feature descriptors (SIFT, SURF, BRISK, 
BRIEF, and FREAK).  These performance gains 
come free in terms of both computational complexity as well as actual 
computing time.

With the framework we introduced three simple methods that are all 
applicable in situations where \emph{Ratio-Match} is currently used.  
\emph{Ratio-Match-Ext} and \emph{Mirror-Match} 
can readily and easily replace \emph{Ratio-Match} without any performance penalty 
and with an increase in precision.  However, both are limited to matching 
image pairs like \emph{Ratio-Match}. \emph{Self-Match} on the other hand is on par with \emph{Ratio-Match} across all 
evaluations, but can be applied when matching multiple overlapping images.

\balance
\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\end{document}